### 2023-09-27 supplementals/manuscript_supplementals.Rmd
supplemental materials

---
title: |
  | Supplemental Materials for
  | Quantifying Stability and Change in Personal Culture Using Panel Data
author: 
- 
output:
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    toc: true
geometry: margin=1in 
mainfont: Times New Roman
fontsize: 12pt
header-includes:
  \usepackage{fontspec}
  \usepackage{longtable}
  \usepackage{multirow}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{dcolumn}
  \usepackage{lscape}
  \usepackage{caption}
  \newcommand{\blandscape}{\begin{landscape}}
  \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.align = "center")
```

\newpage
\blandscape

# Supplemental Materials A

In Table A1, we present the survey items, corresponding labels and topic indicators, as well as the variances calculated from our estimations: V(C) refers to intraindividual change, V(D) refers to interindividual differences, and V(R) refers to residual variance. We also present the $\omega$ values.

```{=latex}
\begin{scriptsize}
\begin{longtable}{|p{1.75in}|p{0.3in}|p{0.3in}|p{0.3in}|p{0.3in}|p{2.5in}|p{2.5in}}
\caption*{Table A1: The Descriptives of Variables Used}
\\
\toprule
\multicolumn{1}{l}{} & \emph{\textbf{ω}} & \textbf{V(C)} & \textbf{V(D)} & \textbf{V(R)} & \textbf{Label} & \textbf{Topic} \\ 
\midrule
\multicolumn{7}{l}{\textbf{BHPS}} \\ 
\midrule
cohadv & 0.09 & 0.05 & 0.53 & 0.42 & Advantages In Living As Couple & Gender \& Family \\ 
cohdis & 0.00 & 0.00 & 0.57 & 0.43 & Disadvantages In Living As Couple & Gender \& Family \\ 
crwora & 0.06 & 0.04 & 0.60 & 0.37 & Worry About Being Affected By Crime & Social Life, Social Cohesion \& Trust \\ 
crworb & 0.10 & 0.07 & 0.61 & 0.32 & Extent Of Concern About Crime & Social Life, Social Cohesion \& Trust \\ 
dfwld & 0.03 & 0.02 & 0.54 & 0.44 & Different World In Future For Young & Gender \& Family \\ 
lfimpa & 0.13 & 0.08 & 0.59 & 0.32 & Importance Of: Health & Health \& Morale \\ 
lfimpb & 0.19 & 0.16 & 0.67 & 0.16 & Importance Of: Money & Health \& Morale \\ 
lfimpc & 0.10 & 0.09 & 0.79 & 0.13 & Importance Of: Having Children & Health \& Morale \\ 
lfimpd & 0.20 & 0.15 & 0.60 & 0.25 & Importance Of: Having A Good Job & Health \& Morale \\ 
lfimpe & 0.16 & 0.12 & 0.63 & 0.25 & Importance Of: Being Independent & Health \& Morale \\ 
lfimpf & 0.19 & 0.17 & 0.70 & 0.13 & Importance Of: Owning Own Home & Health \& Morale \\ 
lfimpg & 0.14 & 0.11 & 0.69 & 0.19 & Importance Of: Good Partnership & Health \& Morale \\ 
lfimph & 0.20 & 0.16 & 0.66 & 0.17 & Importance Of: Good Friends & Health \& Morale \\ 
opchd1\_1 & 0.02 & 0.01 & 0.48 & 0.51 & Quality Best Prep. Child: Well Liked & Gender \& Family \\ 
opchd1\_2 & 0.02 & 0.01 & 0.61 & 0.38 & Quality Best Prep. Child: Think For Self & Gender \& Family \\ 
opchd1\_3 & 0.14 & 0.09 & 0.53 & 0.38 & Quality Best Prep. Child: Work Hard & Gender \& Family \\ 
opchd1\_4 & 0.00 & 0.00 & 0.54 & 0.46 & Quality Best Prep. Child: Help Others & Gender \& Family \\ 
opchd1\_5 & 0.01 & 0.00 & 0.58 & 0.41 & Quality Best Prep. Child: Obey Parents & Gender \& Family \\ 
opcls1 & 0.16 & 0.11 & 0.55 & 0.34 & Sees Self As Belonging To Social Class & Subjective Ses \\ 
opcls2 & 0.10 & 0.08 & 0.73 & 0.19 & Subjective Social Class Membership & Subjective Ses \\ 
opcls3 & 0.16 & 0.10 & 0.52 & 0.38 & Social Class Affects Opportunities & Subjective Ses \\ 
opeur1 & 0.13 & 0.11 & 0.70 & 0.19 & Opinion: Uk Membership Of Eu & Politics, Government \& Economy \\ 
opeur2 & 0.13 & 0.10 & 0.71 & 0.19 & Uk Benefited From Being In Eu & Politics, Government \& Economy \\ 
opeur3 & 0.05 & 0.03 & 0.65 & 0.32 & Uk Long Term Policy Re. Eu Should Be & Politics, Government \& Economy \\ 
opfama & 0.10 & 0.07 & 0.63 & 0.30 & Pre-School Child Suffers If Mother Works & Gender \& Family \\ 
opfamb & 0.09 & 0.07 & 0.64 & 0.29 & Family Suffers If Mother Works Full-Time & Gender \& Family \\ 
opfamc & 0.14 & 0.07 & 0.46 & 0.47 & Woman And Family Happier If She Works & Gender \& Family \\ 
opfamd & 0.12 & 0.07 & 0.54 & 0.39 & Husband And Wife Should Both Contribute & Gender \& Family \\ 
opfame & 0.12 & 0.07 & 0.53 & 0.40 & Full Time Job Makes Woman Independent & Gender \& Family \\ 
opfamf & 0.07 & 0.05 & 0.68 & 0.27 & Husband Should Earn, Wife Stay At Home & Gender \& Family \\ 
opfamg & 0.12 & 0.07 & 0.52 & 0.42 & Children Need Father As Much As Mother & Gender \& Family \\ 
opfamh & 0.11 & 0.07 & 0.59 & 0.34 & Employers Should Help With Childcare & Gender \& Family \\ 
opfami & 0.07 & 0.05 & 0.66 & 0.29 & Single Parents Are As Good As Couples & Gender \& Family \\ 
opfamj & 0.15 & 0.13 & 0.75 & 0.12 & Cohabiting Is Always Wrong & Gender \& Family \\ 
opfamk & 0.11 & 0.06 & 0.52 & 0.42 & Adult Children Should Care For Parents & Gender \& Family \\ 
opfaml & 0.11 & 0.07 & 0.56 & 0.37 & Divorce Better Than Unhappy Marriage & Gender \& Family \\ 
opfamm & 0.12 & 0.11 & 0.77 & 0.13 & The Man Should Be Head Of Household & Gender \& Family \\ 
opfamn & 0.08 & 0.07 & 0.82 & 0.11 & Bible Gods Word And True & Religion \& Spirituality \\ 
opfamo & 0.07 & 0.06 & 0.76 & 0.19 & Co-Habiting Is Alright & Gender \& Family \\ 
opfamp & 0.11 & 0.08 & 0.64 & 0.28 & Parents Ought Stay Together For Children & Gender \& Family \\ 
opfamq & 0.11 & 0.07 & 0.61 & 0.32 & Marital Status Is Irrelevant To Children & Gender \& Family \\ 
opfamr & 0.05 & 0.04 & 0.76 & 0.19 & Homosexual Relationships Are Wrong & Gender \& Family \\ 
ophla & 0.15 & 0.09 & 0.54 & 0.37 & All Health Care Should Be Free & Politics, Government \& Economy \\ 
ophlb & 0.12 & 0.08 & 0.59 & 0.34 & Compulsory Private Insurance If Can Pay & Politics, Government \& Economy \\ 
ophlc & 0.17 & 0.11 & 0.52 & 0.38 & Unfair That Wealth Buys Medical Priority & Politics, Government \& Economy \\ 
opiss1 & 0.07 & 0.05 & 0.68 & 0.28 & Rising Prices Are A Cause For Concern & Politics, Government \& Economy \\ 
opiss2 & 0.05 & 0.03 & 0.71 & 0.26 & Ozone Layer Is A Cause For Concern & Environment \& Climate \\ 
opiss3 & 0.10 & 0.07 & 0.62 & 0.31 & Unemployment Is A Cause For Concern & Politics, Government \& Economy \\ 
opiss4 & 0.01 & 0.00 & 0.71 & 0.29 & Threat To Nature Is A Cause For Concern & Environment \& Climate \\ 
opiss5 & 0.08 & 0.06 & 0.70 & 0.24 & Moral Standards Are A Cause For Concern & Gender \& Family \\ 
opnata & 0.18 & 0.13 & 0.56 & 0.31 & Britain Has Much To Learn & Politics, Government \& Economy \\ 
opnatb & 0.14 & 0.11 & 0.63 & 0.26 & British Citizenship Best & Politics, Government \& Economy \\ 
opnatc & 0.15 & 0.10 & 0.55 & 0.36 & Some Shameful Things About Britain & Politics, Government \& Economy \\ 
opnatd & 0.24 & 0.17 & 0.52 & 0.32 & People Too Ready To Criticise Britain & Politics, Government \& Economy \\ 
opnate & 0.17 & 0.12 & 0.60 & 0.28 & Britain Should Stay Single State & Politics, Government \& Economy \\ 
opnatf & 0.12 & 0.08 & 0.62 & 0.30 & Co-Operation Better Than Independence & Politics, Government \& Economy \\ 
opngbha & 0.17 & 0.13 & 0.65 & 0.22 & Belong To Neighbourhood & Social Life, Social Cohesion \& Trust \\ 
opngbhb & 0.20 & 0.16 & 0.66 & 0.18 & Local Friends Mean A Lot & Social Life, Social Cohesion \& Trust \\ 
opngbhc & 0.15 & 0.11 & 0.63 & 0.26 & Advice Obtainable Locally & Social Life, Social Cohesion \& Trust \\ 
opngbhd & 0.21 & 0.16 & 0.63 & 0.21 & Can Borrow Things From Neighbours & Social Life, Social Cohesion \& Trust \\ 
opngbhe & 0.20 & 0.15 & 0.59 & 0.26 & Willing To Improve Neighbourhood & Social Life, Social Cohesion \& Trust \\ 
opngbhf & 0.10 & 0.07 & 0.59 & 0.34 & Plan To Stay In Neighbourhood & Social Life, Social Cohesion \& Trust \\ 
opngbhg & 0.18 & 0.14 & 0.66 & 0.20 & Am Similar To Others In Neighbourhood & Social Life, Social Cohesion \& Trust \\ 
opngbhh & 0.19 & 0.15 & 0.65 & 0.20 & Talk Regularly To Neighbours & Social Life, Social Cohesion \& Trust \\ 
oppola & 0.19 & 0.11 & 0.48 & 0.41 & Government Reflects People's Wishes & Politics, Government \& Economy \\ 
oppolb & 0.17 & 0.09 & 0.47 & 0.44 & People Can't Influence Govt. Policy & Politics, Government \& Economy \\ 
oppolc & 0.06 & 0.04 & 0.60 & 0.36 & Should Be Max. Limit On Income & Politics, Government \& Economy \\ 
oppold & 0.18 & 0.10 & 0.47 & 0.43 & Govt. Puts Nation's Interests First & Politics, Government \& Economy \\ 
oprlg3 & 0.08 & 0.06 & 0.76 & 0.17 & Religion Makes Difference To Life & Religion \& Spirituality \\ 
opsoca & 0.13 & 0.08 & 0.54 & 0.38 & Ordinary People Share Nations Wealth & Politics, Government \& Economy \\ 
opsocb & 0.11 & 0.07 & 0.58 & 0.34 & One Law For Rich And One For Poor & Politics, Government \& Economy \\ 
opsocc & 0.09 & 0.06 & 0.58 & 0.36 & Private Enterprise Solves Economic Probs & Politics, Government \& Economy \\ 
opsocd & 0.12 & 0.07 & 0.53 & 0.40 & Public Services Ought To Be State Owned & Politics, Government \& Economy \\ 
opsoce & 0.10 & 0.07 & 0.64 & 0.29 & Govt. Has Obligation To Provide Jobs & Politics, Government \& Economy \\ 
opsocf & 0.09 & 0.06 & 0.65 & 0.28 & Strong Trade Unions Protect Employees & Politics, Government \& Economy \\ 
trust & 0.12 & 0.08 & 0.61 & 0.31 & Trustworthiness Of Others & Social Life, Social Cohesion \& Trust \\ 
vote1 & 0.10 & 0.06 & 0.54 & 0.40 & Supports A Particular Political Party & Politics, Government \& Economy \\ 
vote2 & 0.14 & 0.07 & 0.44 & 0.50 & Closer To One Political Party Than Other & Politics, Government \& Economy \\ 
vote4\_1 & 0.05 & 0.04 & 0.88 & 0.08 & Party Affinity: Conservative & Politics, Government \& Economy \\ 
vote4\_2 & 0.05 & 0.04 & 0.85 & 0.11 & Party Affinity: Labour & Politics, Government \& Economy \\ 
vote4\_3 & 0.08 & 0.07 & 0.72 & 0.21 & Party Affinity: Liberal Democrat/Lib/Sdp & Politics, Government \& Economy \\ 
vote5 & 0.10 & 0.06 & 0.54 & 0.41 & Strength Of Support For Stated Party & Politics, Government \& Economy \\ 
vote6 & 0.06 & 0.04 & 0.70 & 0.25 & Level Of Interest In Politics & Politics, Government \& Economy \\ 
\midrule
\multicolumn{7}{l}{\textbf{GSS}} \\ 
\midrule
abany & 0.03 & 0.02 & 0.76 & 0.22 & Abortion: Any Reason & Gender \& Family \\ 
abdefect & 0.10 & 0.08 & 0.75 & 0.17 & Abortion: Birth Defect & Gender \& Family \\ 
abhlth & 0.14 & 0.12 & 0.72 & 0.16 & Abortion: Mother's Health & Gender \& Family \\ 
abnomore & 0.10 & 0.08 & 0.76 & 0.16 & Abortion: Birth Control & Gender \& Family \\ 
abpoor & 0.09 & 0.08 & 0.77 & 0.15 & Abortion: Too Poor & Gender \& Family \\ 
abrape & 0.12 & 0.11 & 0.77 & 0.13 & Abortion: Rape & Gender \& Family \\ 
abscale6 & 0.08 & 0.07 & 0.84 & 0.09 & Abortion Scale (6 Items) & Gender \& Family \\ 
abscale7 & 0.07 & 0.06 & 0.84 & 0.10 & Abortion Scale (7 Items) & Gender \& Family \\ 
absingle & 0.05 & 0.04 & 0.77 & 0.19 & Abortion: Unmarried & Gender \& Family \\ 
affrmact & 0.12 & 0.09 & 0.68 & 0.22 & Hiring Preferences & National Identity, Ethnicity \& Immigration \\ 
aged & 0.01 & 0.00 & 0.60 & 0.40 & Seniors Should Live With Family & Gender \& Family \\ 
athscale & 0.00 & 0.00 & 0.74 & 0.26 & Athiest: Scale & Politics, Government \& Economy \\ 
attend & 0.08 & 0.08 & 0.85 & 0.07 & Attend Religious Services & Religion \& Spirituality \\ 
bible & 0.10 & 0.09 & 0.80 & 0.12 & View Of Bible & Religion \& Spirituality \\ 
cappun & 0.12 & 0.11 & 0.76 & 0.13 & Death Penalty & Politics, Government \& Economy \\ 
chldidel & 0.15 & 0.12 & 0.71 & 0.17 & Ideal \# Of Children & Gender \& Family \\ 
class & 0.09 & 0.07 & 0.73 & 0.20 & Social Class & Subjective Ses \\ 
closeblk & 0.11 & 0.09 & 0.71 & 0.20 & Feel Close To Blacks & National Identity, Ethnicity \& Immigration \\ 
closewht & 0.04 & 0.03 & 0.61 & 0.36 & Feel Close To Whites & National Identity, Ethnicity \& Immigration \\ 
colath & 0.00 & 0.00 & 0.64 & 0.36 & Athiest: Teach At College & Politics, Government \& Economy \\ 
colcom & 0.09 & 0.06 & 0.63 & 0.31 & Red: Teach At College & Politics, Government \& Economy \\ 
colhomo & 0.00 & 0.00 & 0.67 & 0.32 & Gay: Teach At College & Politics, Government \& Economy \\ 
colmil & 0.03 & 0.02 & 0.61 & 0.37 & Militarist: Teach At College & Politics, Government \& Economy \\ 
colmslm & 0.00 & 0.00 & 0.70 & 0.30 & Islamist: Teach At College & Politics, Government \& Economy \\ 
colrac & 0.00 & 0.00 & 0.60 & 0.40 & Racist: Teach At College & Politics, Government \& Economy \\ 
comscale & 0.03 & 0.02 & 0.76 & 0.22 & Red: Scale & Politics, Government \& Economy \\ 
conarmy & 0.12 & 0.08 & 0.63 & 0.29 & Military & Social Life, Social Cohesion \& Trust \\ 
conbus & 0.11 & 0.07 & 0.60 & 0.33 & Major Companies & Social Life, Social Cohesion \& Trust \\ 
conclerg & 0.05 & 0.03 & 0.66 & 0.31 & Organized Religion & Social Life, Social Cohesion \& Trust \\ 
coneduc & 0.02 & 0.01 & 0.60 & 0.38 & Education & Social Life, Social Cohesion \& Trust \\ 
confed & 0.24 & 0.17 & 0.56 & 0.27 & Executive Branch & Social Life, Social Cohesion \& Trust \\ 
confinan & 0.10 & 0.06 & 0.56 & 0.38 & Banks \& Finance & Social Life, Social Cohesion \& Trust \\ 
conjudge & 0.09 & 0.06 & 0.63 & 0.31 & Supreme Court & Social Life, Social Cohesion \& Trust \\ 
conlabor & 0.14 & 0.10 & 0.62 & 0.28 & Organized Labor & Social Life, Social Cohesion \& Trust \\ 
conlegis & 0.02 & 0.01 & 0.61 & 0.37 & Congress & Social Life, Social Cohesion \& Trust \\ 
conmedia & 0.17 & 0.13 & 0.66 & 0.21 & Tv \& Press & Social Life, Social Cohesion \& Trust \\ 
conmedic & 0.13 & 0.09 & 0.59 & 0.32 & Medicine & Social Life, Social Cohesion \& Trust \\ 
conpress & 0.16 & 0.12 & 0.62 & 0.26 & Press & Social Life, Social Cohesion \& Trust \\ 
consci & 0.04 & 0.02 & 0.61 & 0.36 & Scientific Community & Social Life, Social Cohesion \& Trust \\ 
contv & 0.18 & 0.13 & 0.59 & 0.27 & Television & Social Life, Social Cohesion \& Trust \\ 
courts & 0.06 & 0.04 & 0.65 & 0.32 & Courts Deal With Criminals & Politics, Government \& Economy \\ 
discaff & 0.00 & 0.00 & 0.56 & 0.44 & Affirm. Action Hurt Whites & National Identity, Ethnicity \& Immigration \\ 
discaffm & 0.09 & 0.06 & 0.55 & 0.40 & Affirm. Action Hurt Men & Gender \& Family \\ 
discaffw & 0.03 & 0.02 & 0.56 & 0.42 & Affirm. Action Hurt Women & Gender \& Family \\ 
divlaw & 0.00 & 0.00 & 0.71 & 0.29 & Divorce Laws Too Lenient & Politics, Government \& Economy \\ 
eqwlth & 0.05 & 0.04 & 0.68 & 0.28 & Government Reduce Inequality & Politics, Government \& Economy \\ 
fair & 0.04 & 0.03 & 0.68 & 0.30 & People Try To Be Fair & Social Life, Social Cohesion \& Trust \\ 
fear & 0.06 & 0.04 & 0.66 & 0.30 & Afraid To Walk At Night & Politics, Government \& Economy \\ 
fechld & 0.13 & 0.10 & 0.65 & 0.25 & Working Moms Warm & Gender \& Family \\ 
fefam & 0.02 & 0.01 & 0.69 & 0.30 & Better For Man To Work & Gender \& Family \\ 
fehire & 0.00 & 0.00 & 0.62 & 0.38 & Hire \& Promote Women & Gender \& Family \\ 
fejobaff & 0.16 & 0.13 & 0.69 & 0.18 & Preferences For Hiring Women & Gender \& Family \\ 
fepol & 0.00 & 0.00 & 0.64 & 0.36 & Men Better For Politics & Gender \& Family \\ 
fepresch & 0.02 & 0.01 & 0.66 & 0.33 & Pre-Ks Suffer If Mom Works & Gender \& Family \\ 
finalter & 0.20 & 0.14 & 0.54 & 0.32 & Finances Better Or Worse & Subjective Ses \\ 
finrela & 0.10 & 0.08 & 0.70 & 0.22 & Relative Income & Subjective Ses \\ 
fund & 0.08 & 0.07 & 0.84 & 0.09 & Fundamentalist & Religion \& Spirituality \\ 
genderscale & 0.03 & 0.02 & 0.76 & 0.22 & Scale Of Gender Role Attitudes & Gender \& Family \\ 
getahead & 0.19 & 0.13 & 0.56 & 0.31 & Work To Get Ahead & Subjective Ses \\ 
god & 0.05 & 0.04 & 0.82 & 0.14 & Belief About God & Religion \& Spirituality \\ 
goodlife & 0.05 & 0.03 & 0.60 & 0.37 & Finances Will Improve & Subjective Ses \\ 
grass & 0.03 & 0.03 & 0.78 & 0.20 & Legalize Marijuana & Politics, Government \& Economy \\ 
gunlaw & 0.00 & 0.00 & 0.63 & 0.37 & Require Gun Permits & Politics, Government \& Economy \\ 
hapmar & 0.08 & 0.06 & 0.70 & 0.24 & Happiness: Marriage & Health \& Morale \\ 
happy & 0.19 & 0.15 & 0.62 & 0.23 & Happiness: General & Health \& Morale \\ 
health & 0.10 & 0.08 & 0.75 & 0.17 & Health: Self-Rated & Health \& Morale \\ 
helpblk & 0.04 & 0.03 & 0.70 & 0.27 & Government Help Blacks & Politics, Government \& Economy \\ 
helpful & 0.01 & 0.01 & 0.65 & 0.34 & People Try To Be Helpful & Social Life, Social Cohesion \& Trust \\ 
helpnot & 0.00 & 0.00 & 0.62 & 0.37 & Government Do More Or Less & Politics, Government \& Economy \\ 
helpoth & 0.10 & 0.07 & 0.59 & 0.34 & Child Should Help Others & Gender \& Family \\ 
helppoor & 0.01 & 0.01 & 0.65 & 0.35 & Government Help Poor & Politics, Government \& Economy \\ 
helpsick & 0.02 & 0.01 & 0.66 & 0.33 & Government Pay Medical & Politics, Government \& Economy \\ 
homoscale & 0.00 & 0.00 & 0.74 & 0.26 & Gay: Scale & Politics, Government \& Economy \\ 
homosex & 0.04 & 0.04 & 0.83 & 0.13 & Morality: Same-Sex Sex & Gender \& Family \\ 
intlblks & 0.23 & 0.14 & 0.48 & 0.37 & Blacks Intelligent Or Not & National Identity, Ethnicity \& Immigration \\ 
intlwhts & 0.01 & 0.01 & 0.49 & 0.50 & Whites Intelligent Or Not & National Identity, Ethnicity \& Immigration \\ 
jobfind & 0.16 & 0.12 & 0.65 & 0.23 & Find A Good Job & Subjective Ses \\ 
joblose & 0.01 & 0.00 & 0.56 & 0.44 & Likely To Lose Job & Subjective Ses \\ 
kidssol & 0.04 & 0.03 & 0.69 & 0.28 & Mobility Next Generation & Subjective Ses \\ 
letdie1 & 0.14 & 0.12 & 0.72 & 0.17 & Let Die: Incurable & Health \& Morale \\ 
letin1a & 0.02 & 0.01 & 0.70 & 0.29 & How Many Immigrants? & National Identity, Ethnicity \& Immigration \\ 
libath & 0.00 & 0.00 & 0.61 & 0.39 & Athiest: Library Book & Politics, Government \& Economy \\ 
libcom & 0.05 & 0.04 & 0.65 & 0.31 & Red: Library Book & Politics, Government \& Economy \\ 
libhomo & 0.00 & 0.00 & 0.65 & 0.35 & Gay: Library Book & Politics, Government \& Economy \\ 
libmil & 0.02 & 0.01 & 0.60 & 0.38 & Militarist: Library Book & Politics, Government \& Economy \\ 
libmslm & 0.00 & 0.00 & 0.68 & 0.32 & Islamist: Library Book & Politics, Government \& Economy \\ 
librac & 0.00 & 0.00 & 0.59 & 0.40 & Racist: Library Book & Politics, Government \& Economy \\ 
life & 0.04 & 0.03 & 0.67 & 0.30 & Life: Exciting Or Dull & Health \& Morale \\ 
liveblks & 0.20 & 0.14 & 0.58 & 0.27 & Neighborhood Half Black & National Identity, Ethnicity \& Immigration \\ 
livewhts & 0.13 & 0.07 & 0.49 & 0.44 & Neighborhood Half White & National Identity, Ethnicity \& Immigration \\ 
marblk & 0.06 & 0.04 & 0.69 & 0.27 & Relative Marry Black & National Identity, Ethnicity \& Immigration \\ 
marhomo & 0.05 & 0.04 & 0.82 & 0.14 & Gay Marriage & Gender \& Family \\ 
marwht & 0.00 & 0.00 & 0.55 & 0.45 & Relative Marry White & National Identity, Ethnicity \& Immigration \\ 
meovrwrk & 0.03 & 0.02 & 0.58 & 0.40 & Men Hurt Family When Overwork & Gender \& Family \\ 
milscale & 0.03 & 0.02 & 0.69 & 0.29 & Militarist: Scale & Politics, Government \& Economy \\ 
misanthropy & 0.07 & 0.06 & 0.77 & 0.17 & Misanthropy Scale & Social Life, Social Cohesion \& Trust \\ 
mslmscale & 0.00 & 0.00 & 0.77 & 0.23 & Islamist: Scale & Politics, Government \& Economy \\ 
nataid & 0.07 & 0.05 & 0.65 & 0.30 & Foreign Aid & Politics, Government \& Economy \\ 
natarms & 0.08 & 0.06 & 0.67 & 0.28 & Defense & Politics, Government \& Economy \\ 
natchld & 0.05 & 0.04 & 0.63 & 0.34 & Child Care & Politics, Government \& Economy \\ 
natcity & 0.05 & 0.03 & 0.63 & 0.34 & Problems Of Cities & Politics, Government \& Economy \\ 
natcrime & 0.12 & 0.08 & 0.62 & 0.29 & Fighting Crime & Politics, Government \& Economy \\ 
natdrug & 0.12 & 0.08 & 0.60 & 0.32 & Dealing With Drugs & Politics, Government \& Economy \\ 
nateduc & 0.16 & 0.13 & 0.66 & 0.21 & Education & Politics, Government \& Economy \\ 
natenvir & 0.01 & 0.01 & 0.67 & 0.32 & Protect Environment & Politics, Government \& Economy \\ 
natfare & 0.08 & 0.06 & 0.75 & 0.19 & Welfare & Politics, Government \& Economy \\ 
natheal & 0.02 & 0.01 & 0.57 & 0.41 & Improving Health & Politics, Government \& Economy \\ 
natmass & 0.05 & 0.03 & 0.62 & 0.35 & Mass Transit & Politics, Government \& Economy \\ 
natpark & 0.19 & 0.14 & 0.58 & 0.28 & Parks \& Recreation & Politics, Government \& Economy \\ 
natrace & 0.13 & 0.11 & 0.72 & 0.18 & Improve Life For Blacks & Politics, Government \& Economy \\ 
natroad & 0.01 & 0.01 & 0.61 & 0.39 & Highways \& Bridges & Politics, Government \& Economy \\ 
natsci & 0.12 & 0.08 & 0.60 & 0.32 & Scientific Research & Politics, Government \& Economy \\ 
natsoc & 0.01 & 0.00 & 0.63 & 0.36 & Social Security & Politics, Government \& Economy \\ 
natspac & 0.10 & 0.07 & 0.70 & 0.23 & Space Exploration & Politics, Government \& Economy \\ 
neverrelact & 0.12 & 0.09 & 0.65 & 0.27 & Never Active In Religion & Religion \& Spirituality \\ 
news & 0.08 & 0.06 & 0.77 & 0.16 & Read Newspaper & Social Life, Social Cohesion \& Trust \\ 
obey & 0.08 & 0.07 & 0.72 & 0.22 & Child Should Obey & Gender \& Family \\ 
owngun & 0.08 & 0.08 & 0.86 & 0.06 & Gun Owner & Politics, Government \& Economy \\ 
parsol & 0.01 & 0.01 & 0.71 & 0.28 & Mobility & Subjective Ses \\ 
partyid & 0.07 & 0.06 & 0.88 & 0.05 & Political Party & Politics, Government \& Economy \\ 
pillok & 0.10 & 0.07 & 0.70 & 0.23 & Birth Control 14-16 Year Olds & Gender \& Family \\ 
polabuse & 0.00 & 0.00 & 0.50 & 0.50 & Police Hit Person Cursing & Politics, Government \& Economy \\ 
polattak & 0.00 & 0.00 & 0.54 & 0.46 & Police Hit Person Punching & Politics, Government \& Economy \\ 
polescap & 0.16 & 0.11 & 0.59 & 0.30 & Police Hit Person Fleeing & Politics, Government \& Economy \\ 
polhitok & 0.00 & 0.00 & 0.71 & 0.29 & Police Can Hit Citizens & Politics, Government \& Economy \\ 
polmurdr & 0.00 & 0.00 & 0.53 & 0.47 & Police Hit Murder Suspect & Politics, Government \& Economy \\ 
polviews & 0.08 & 0.07 & 0.75 & 0.18 & Political Views & Politics, Government \& Economy \\ 
polviolent & 0.01 & 0.01 & 0.62 & 0.38 & Police Hit: Scale & Politics, Government \& Economy \\ 
popespks & 0.15 & 0.12 & 0.70 & 0.19 & Pope Infallible & Religion \& Spirituality \\ 
popular & 0.23 & 0.17 & 0.55 & 0.28 & Child Should Be Popular & Gender \& Family \\ 
pornlaw & 0.04 & 0.02 & 0.66 & 0.32 & Restrict Pornography & Gender \& Family \\ 
postlife & 0.09 & 0.07 & 0.78 & 0.15 & Life After Death & Religion \& Spirituality \\ 
pray & 0.08 & 0.07 & 0.86 & 0.07 & Pray & Religion \& Spirituality \\ 
prayer & 0.13 & 0.10 & 0.68 & 0.22 & Bible Prayer In Schools & Politics, Government \& Economy \\ 
premarsx & 0.07 & 0.06 & 0.79 & 0.15 & Morality: Premarital Sex & Gender \& Family \\ 
racdif1 & 0.09 & 0.07 & 0.65 & 0.28 & Racial Diff.: Discrimination & National Identity, Ethnicity \& Immigration \\ 
racdif2 & 0.00 & 0.00 & 0.57 & 0.43 & Racial Diff.: Inborn & National Identity, Ethnicity \& Immigration \\ 
racdif3 & 0.08 & 0.05 & 0.64 & 0.30 & Racial Diff.: Lack Education & National Identity, Ethnicity \& Immigration \\ 
racdif4 & 0.06 & 0.04 & 0.64 & 0.32 & Racial Diff.: Lack Will & National Identity, Ethnicity \& Immigration \\ 
racopen & 0.13 & 0.09 & 0.61 & 0.30 & Open Housing Law & National Identity, Ethnicity \& Immigration \\ 
racscale & 0.01 & 0.01 & 0.69 & 0.31 & Racist: Scale & Politics, Government \& Economy \\ 
reborn & 0.07 & 0.06 & 0.81 & 0.13 & Born Again & Religion \& Spirituality \\ 
relactiv & 0.13 & 0.11 & 0.76 & 0.13 & Do Religious Activities & Religion \& Spirituality \\ 
relexp & 0.14 & 0.12 & 0.76 & 0.12 & Had Religious Experience & Religion \& Spirituality \\ 
relexper & 0.03 & 0.02 & 0.74 & 0.24 & Religious Turning Point & Religion \& Spirituality \\ 
reliten & 0.10 & 0.09 & 0.80 & 0.11 & Strength Of Affiliation & Religion \& Spirituality \\ 
relpersn & 0.06 & 0.05 & 0.81 & 0.14 & Religious Person & Religion \& Spirituality \\ 
richwork & 0.11 & 0.08 & 0.66 & 0.26 & Would Work If Rich & Health \& Morale \\ 
satfin & 0.17 & 0.14 & 0.68 & 0.18 & Satisfied With Finances & Health \& Morale \\ 
satjob & 0.20 & 0.15 & 0.60 & 0.25 & Satisfied With Work & Health \& Morale \\ 
selfhunt & 0.02 & 0.02 & 0.80 & 0.18 & Hunter: Self & Politics, Government \& Economy \\ 
sexeduc & 0.00 & 0.00 & 0.69 & 0.31 & Sex-Ed. In Public Schools & Gender \& Family \\ 
socbar & 0.11 & 0.10 & 0.80 & 0.10 & Meet Friends At Bar & Social Life, Social Cohesion \& Trust \\ 
socfrend & 0.09 & 0.06 & 0.65 & 0.28 & Meet Friends & Social Life, Social Cohesion \& Trust \\ 
socialscale & 0.09 & 0.07 & 0.74 & 0.18 & Social Life Scale & Social Life, Social Cohesion \& Trust \\ 
socommun & 0.16 & 0.12 & 0.64 & 0.25 & Meet Neighbors & Social Life, Social Cohesion \& Trust \\ 
socrel & 0.13 & 0.10 & 0.67 & 0.23 & Meet Relatives & Social Life, Social Cohesion \& Trust \\ 
spanking & 0.07 & 0.06 & 0.73 & 0.21 & Spank Children & Gender \& Family \\ 
spkath & 0.03 & 0.02 & 0.66 & 0.32 & Athiest: Make Speech & Politics, Government \& Economy \\ 
spkcom & 0.01 & 0.01 & 0.71 & 0.28 & Red: Make Speech & Politics, Government \& Economy \\ 
spkhomo & 0.07 & 0.05 & 0.65 & 0.30 & Gay: Make Speech & Politics, Government \& Economy \\ 
spkmil & 0.00 & 0.00 & 0.63 & 0.36 & Militarist: Make Speech & Politics, Government \& Economy \\ 
spkmslm & 0.02 & 0.01 & 0.73 & 0.26 & Islamist: Make Speech & Politics, Government \& Economy \\ 
spkrac & 0.06 & 0.04 & 0.63 & 0.33 & Racist: Make Speech & Politics, Government \& Economy \\ 
sprtprsn & 0.10 & 0.09 & 0.78 & 0.13 & Spiritual Person & Religion \& Spirituality \\ 
suicide1 & 0.01 & 0.00 & 0.73 & 0.27 & Suicide: Incurable & Health \& Morale \\ 
suicide2 & 0.08 & 0.06 & 0.65 & 0.29 & Suicide: Bankruptcy & Health \& Morale \\ 
suicide3 & 0.03 & 0.02 & 0.66 & 0.32 & Suicide: Dishonor & Health \& Morale \\ 
suicide4 & 0.08 & 0.06 & 0.68 & 0.26 & Suicide: Tired Of Living & Health \& Morale \\ 
suicidescale & 0.03 & 0.03 & 0.75 & 0.23 & Suicide Scale & Health \& Morale \\ 
teensex & 0.11 & 0.08 & 0.66 & 0.26 & Morality: Sex 14-16 Year Olds & Gender \& Family \\ 
thnkself & 0.10 & 0.07 & 0.65 & 0.27 & Child Should Think For Self & Gender \& Family \\ 
trust & 0.00 & 0.00 & 0.72 & 0.28 & People Can Be Trusted & Social Life, Social Cohesion \& Trust \\ 
tvhours & 0.15 & 0.13 & 0.75 & 0.13 & Watch Television & Social Life, Social Cohesion \& Trust \\ 
uswary & 0.04 & 0.03 & 0.67 & 0.30 & Us At War Next 10 Years & Politics, Government \& Economy \\ 
wlthblks & 0.02 & 0.01 & 0.52 & 0.47 & Blacks Rich Or Poor & National Identity, Ethnicity \& Immigration \\ 
wlthwhts & 0.23 & 0.16 & 0.53 & 0.30 & Whites Rich Or Poor & National Identity, Ethnicity \& Immigration \\ 
workblks & 0.00 & 0.00 & 0.56 & 0.44 & Blacks Work Hard & National Identity, Ethnicity \& Immigration \\ 
workhard & 0.01 & 0.00 & 0.57 & 0.42 & Child Should Work Hard & Gender \& Family \\ 
workwhts & 0.23 & 0.16 & 0.55 & 0.29 & Whites Work Hard & National Identity, Ethnicity \& Immigration \\ 
wrkwayup & 0.12 & 0.09 & 0.69 & 0.21 & No Racial Favors & National Identity, Ethnicity \& Immigration \\ 
xmarsex & 0.18 & 0.14 & 0.64 & 0.22 & Morality: Extramarital Affair & Gender \& Family \\ 
xmovie & 0.07 & 0.06 & 0.71 & 0.23 & Watch X-Rated Movie & Social Life, Social Cohesion \& Trust \\ 
\midrule
\multicolumn{7}{l}{\textbf{HILDA}} \\ 
\midrule
atwkadc & 0.13 & 0.09 & 0.58 & 0.33 & Fine For Child Under 3 To Be In All-Day Care & Gender \& Family \\ 
atwkbmw & 0.07 & 0.05 & 0.69 & 0.25 & Better If Man Earns Money, Woman Cares For Children & Gender \& Family \\ 
atwkcdw & 0.13 & 0.09 & 0.58 & 0.33 & As Well For Child If Mother Earns Money, Father Cares For Children & Gender \& Family \\ 
atwkcs & 0.12 & 0.08 & 0.59 & 0.33 & Children Suffer Bc Father Works Too Much & Gender \& Family \\ 
atwkdcm & 0.14 & 0.10 & 0.64 & 0.25 & Better For Child To Stay With Mother In Case Of Divorce & Gender \& Family \\ 
atwkfhi & 0.17 & 0.11 & 0.55 & 0.33 & Father Should Be As Involved As Mother & Gender \& Family \\ 
atwkmmf & 0.08 & 0.05 & 0.61 & 0.34 & Bad For Relationship If Woman Earns More & Gender \& Family \\ 
atwkmpl & 0.02 & 0.01 & 0.66 & 0.33 & Men Make Better Political Leaders & Gender \& Family \\ 
atwkmrl & 0.11 & 0.08 & 0.59 & 0.34 & Most Important Role For Men To Be Father & Gender \& Family \\ 
atwkmsw & 0.08 & 0.05 & 0.62 & 0.33 & Mothers Who Don't Need Money Shouldn't Work & Gender \& Family \\ 
atwkpsc & 0.08 & 0.06 & 0.63 & 0.32 & Pre-School Child Suffers If Mother Works Ft & Gender \& Family \\ 
atwkseh & 0.16 & 0.10 & 0.53 & 0.37 & If Both Partners Work They Should Share Domestic Tasks & Gender \& Family \\ 
atwkwfr & 0.12 & 0.07 & 0.52 & 0.41 & Working Father Can Have As Good Relationship To Child & Gender \& Family \\ 
atwkwfs & 0.15 & 0.09 & 0.52 & 0.39 & Working Fathers Care More About Success Than Needs Of Children & Gender \& Family \\ 
atwkwmr & 0.11 & 0.07 & 0.58 & 0.36 & Working Mother Can Have As Good Relationship To Child & Gender \& Family \\ 
atwkwms & 0.15 & 0.10 & 0.56 & 0.34 & Working Mothers Care More About Success Than Needs Of Children & Gender \& Family \\ 
atwkwrl & 0.10 & 0.07 & 0.60 & 0.33 & Most Important Role For Women To Be Mother & Gender \& Family \\ 
lskeep & 0.07 & 0.04 & 0.55 & 0.41 & Most People You Meet Keep Their Word & Social Life, Social Cohesion \& Trust \\ 
lssop & 0.13 & 0.08 & 0.58 & 0.34 & Most People You Meet Succeed By Stepping On Others & Social Life, Social Cohesion \& Trust \\ 
mcchmf & 0.12 & 0.09 & 0.66 & 0.25 & Children Happier With Both Father And Mother & Gender \& Family \\ 
mcdef & 0.06 & 0.05 & 0.80 & 0.15 & Cohabitation Alright Even Without Marriage Intention & Gender \& Family \\ 
mcdiv & 0.08 & 0.05 & 0.62 & 0.32 & Divorce Alright Even With Children & Gender \& Family \\ 
mchscr & 0.09 & 0.07 & 0.76 & 0.16 & Same Rights For Homosexual Couples & Gender \& Family \\ 
mcindpc & 0.15 & 0.11 & 0.63 & 0.26 & Child Should Start Living Independently By 18-20 & Gender \& Family \\ 
mcmhc & 0.18 & 0.14 & 0.66 & 0.20 & A Man Has To Have Children In Order To Be Fulfilled & Gender \& Family \\ 
mcmoi & 0.13 & 0.09 & 0.64 & 0.27 & Marriage Is Outdated & Gender \& Family \\ 
mcnodiv & 0.08 & 0.07 & 0.70 & 0.23 & Marriage Should Never Be Ended & Gender \& Family \\ 
mcsingp & 0.07 & 0.05 & 0.69 & 0.26 & Alright For Woman To Have Children As Single Parent & Gender \& Family \\ 
mcwhc & 0.16 & 0.12 & 0.67 & 0.21 & A Woman Has To Have Children In Order To Be Fulfilled & Gender \& Family \\ 
relimp & 0.05 & 0.05 & 0.87 & 0.08 & Importance Of Religion & Religion \& Spirituality \\ 
\midrule
\multicolumn{7}{l}{\textbf{PSID}} \\ 
\midrule
cds\_bothparentscare & 0.06 & 0.04 & 0.55 & 0.42 & Dad Should Be As Involved As Mom & Gender \& Family \\ 
cds\_daddevelopment & 0.08 & 0.05 & 0.53 & 0.42 & Dad Large Role In Child Development & Gender \& Family \\ 
cds\_dadoldkidbetter & 0.15 & 0.11 & 0.59 & 0.30 & Dad Able To Enjoy Older Children More & Gender \& Family \\ 
cds\_dadtooinvolved & 0.19 & 0.13 & 0.55 & 0.31 & Dad Too Involved If Keeps From Better Job & Gender \& Family \\ 
cds\_daycareok & 0.16 & 0.12 & 0.63 & 0.25 & Ok To Have Child Under 3 In All-Day Care & Gender \& Family \\ 
cds\_fatherhoodfulfill & 0.00 & 0.00 & 0.57 & 0.43 & Dadhood Most Fulfilling Experience & Gender \& Family \\ 
cds\_fatherinteractchild & 0.01 & 0.00 & 0.56 & 0.44 & Dads Should Interact With Child & Gender \& Family \\ 
cds\_firstyears & 0.01 & 0.00 & 0.59 & 0.40 & First 4 Years Most Import. In Child's Develop. & Gender \& Family \\ 
cds\_girlboyequal & 0.05 & 0.03 & 0.59 & 0.38 & Encourage As Much Independence In Daughters And Sons & Gender \& Family \\ 
cds\_husbandcareerimport & 0.13 & 0.09 & 0.61 & 0.30 & Better Wife Helps Husband Career & Gender \& Family \\ 
cds\_mostimportantchild\_1 & 0.10 & 0.06 & 0.55 & 0.39 & Importance For Child: Obey & Gender \& Family \\ 
cds\_mostimportantchild\_3 & 0.00 & 0.00 & 0.57 & 0.43 & Importance For Child: Think For Herself & Gender \& Family \\ 
cds\_motherfatherequal & 0.19 & 0.13 & 0.56 & 0.31 & Dad And Mom Equally Good Meeting Kid's Needs & Gender \& Family \\ 
cds\_motherhoodfulfill & 0.00 & 0.00 & 0.59 & 0.41 & Momhood Most Fulfilling Experience & Gender \& Family \\ 
cds\_mothernotft & 0.12 & 0.10 & 0.67 & 0.24 & Mom With Child Under 5 Shouldn't Work Full-Time & Gender \& Family \\ 
cds\_preschoolsuffer & 0.13 & 0.09 & 0.65 & 0.26 & Preschool Child Suffers If Mom Employed & Gender \& Family \\ 
cds\_sharehousehold & 0.00 & 0.00 & 0.48 & 0.52 & Husband And Wife In Both Ft Should Share Household & Gender \& Family \\ 
cds\_traditionalbest & 0.10 & 0.07 & 0.66 & 0.27 & Better If Man Earns Living & Gender \& Family \\ 
cds\_workingmotherssame & 0.05 & 0.03 & 0.55 & 0.42 & Working Mom As Warm As Nonworking Mom & Gender \& Family \\ 
gettingbetter & 0.18 & 0.11 & 0.50 & 0.38 & Life Getting Better & Health \& Morale \\ 
lifesatis & 0.17 & 0.11 & 0.53 & 0.35 & Life Satisfaction & Health \& Morale \\ 
likejobormoney & 0.17 & 0.09 & 0.46 & 0.44 & Satisfaction About Income Important In Job & Occupation \& Education \\ 
opinionimport & 0.17 & 0.12 & 0.56 & 0.32 & How Much Does It Matter What Other People Think About You? & Social Life, Social Cohesion \& Trust \\ 
satisfiedself & 0.22 & 0.14 & 0.50 & 0.36 & Are You More Often Satisfied Or Dissatisfied With Yourself? & Health \& Morale \\ 
savingspending & 0.20 & 0.12 & 0.49 & 0.38 & Spend Or Save Money & Health \& Morale \\ 
tas\_belongcommunity & 0.14 & 0.09 & 0.56 & 0.36 & Freq. Of Feeling Belonging To Communty & Social Life, Social Cohesion \& Trust \\ 
tas\_contributesociety & 0.14 & 0.09 & 0.53 & 0.38 & Freq. Of Feeling Contrib To Society & Social Life, Social Cohesion \& Trust \\ 
tas\_divorcehurts & 0.10 & 0.05 & 0.49 & 0.45 & Divorce Hurts Kids & Gender \& Family \\ 
tas\_divorcesuccess & 0.15 & 0.08 & 0.47 & 0.45 & Child From Divorced Family Can Be Successful & Gender \& Family \\ 
tas\_fathertimeoff & 0.20 & 0.13 & 0.52 & 0.36 & Dads Should Take Time With Young Child & Gender \& Family \\ 
tas\_importrel & 0.10 & 0.08 & 0.73 & 0.19 & Importance Of Religion & Religion \& Spirituality \\ 
tas\_importspirit & 0.13 & 0.10 & 0.67 & 0.24 & Importance Of Spirituality & Religion \& Spirituality \\ 
tas\_lifesatisfact & 0.19 & 0.14 & 0.61 & 0.25 & How Satisfied With Life As A Whole & Health \& Morale \\ 
tas\_mothersnodemanding & 0.17 & 0.11 & 0.55 & 0.34 & Moms Should Have Demandng Job & Gender \& Family \\ 
tas\_peoplearegood & 0.15 & 0.10 & 0.54 & 0.36 & Freq. Of Feeling People Basically Good & Social Life, Social Cohesion \& Trust \\ 
tas\_religpref0509\_2 & 0.04 & 0.03 & 0.76 & 0.21 & Religion: Protestant [2005-09] & Religion \& Spirituality \\ 
tas\_religpref0509\_3 & 0.07 & 0.06 & 0.89 & 0.05 & Religion: Catholic [2005-09] & Religion \& Spirituality \\ 
tas\_religpref1115\_2 & 0.11 & 0.10 & 0.83 & 0.07 & Religion: Protestant [2011-15] & Religion \& Spirituality \\ 
tas\_religpref1115\_3 & 0.03 & 0.02 & 0.87 & 0.11 & Religion: Catholic [2011-15] & Religion \& Spirituality \\ 
tas\_singleparentsok & 0.15 & 0.09 & 0.50 & 0.41 & Child From Single-Parent Family Can Do As Well & Gender \& Family \\ 
tas\_singleparentsuffer & 0.14 & 0.10 & 0.60 & 0.30 & Single Parent Hurts Kids & Gender \& Family \\ 
tas\_singlewomenkids & 0.12 & 0.07 & 0.51 & 0.42 & Single Women Should Have Kids & Gender \& Family \\ 
tas\_societybetter & 0.17 & 0.11 & 0.52 & 0.37 & Freq. Of Feeling Society Gettng Better & Social Life, Social Cohesion \& Trust \\ 
tas\_societymakessense & 0.14 & 0.09 & 0.53 & 0.38 & Feeling Way Society Works Makes Sense & Social Life, Social Cohesion \& Trust \\ 
tas\_spritual & 0.14 & 0.10 & 0.63 & 0.27 & Spiritual Person & Religion \& Spirituality \\ 
thinkfuture & 0.21 & 0.13 & 0.50 & 0.37 & Think About Future & Health \& Morale \\ 
trustothers & 0.13 & 0.10 & 0.65 & 0.26 & Trust Most Others & Social Life, Social Cohesion \& Trust \\ 
undeserved & 0.19 & 0.13 & 0.55 & 0.32 & Are There A Lot Of People Who Have Good Things They Don't Deserve? & Social Life, Social Cohesion \& Trust \\ 
\midrule
\multicolumn{7}{l}{\textbf{SHP}} \\ 
\midrule
pc70 & 0.05 & 0.03 & 0.56 & 0.40 & Self-Perception : Incapacity To Make Plans Because Of Unpredictability & Health \& Morale \\ 
pc71 & 0.16 & 0.10 & 0.55 & 0.35 & Self-Perception : Little Influence On Life Events & Health \& Morale \\ 
pc72 & 0.00 & 0.00 & 0.54 & 0.46 & Self-Perception : Capacity To Overcome Unexpected Problems & Health \& Morale \\ 
pc73 & 0.03 & 0.02 & 0.59 & 0.39 & Self-Perception : Capacity To Choose Between Two Possibilities & Health \& Morale \\ 
pc74 & 0.13 & 0.09 & 0.63 & 0.27 & Self-Perception : Feeling Of Uselessness & Health \& Morale \\ 
pc75 & 0.00 & 0.00 & 0.64 & 0.36 & Self-Perception : Feeling Of Selfsatisfaction & Health \& Morale \\ 
pl101 & 0.14 & 0.10 & 0.63 & 0.27 & Feeling Of Security (Victimization) & Health \& Morale \\ 
pn56 & 0.21 & 0.17 & 0.63 & 0.20 & Values: Lying In Own Interest & Social Life, Social Cohesion \& Trust \\ 
pn57 & 0.20 & 0.15 & 0.63 & 0.22 & Values: Cheating On Tax Declaration & Social Life, Social Cohesion \& Trust \\ 
pn58 & 0.20 & 0.16 & 0.64 & 0.20 & Values: Avoiding A Fare On A Public Transport & Social Life, Social Cohesion \& Trust \\ 
pn59 & 0.05 & 0.02 & 0.50 & 0.47 & Values: Claiming State Benefits Not Entitled To & Social Life, Social Cohesion \& Trust \\ 
pn60 & 0.11 & 0.08 & 0.67 & 0.25 & Values: Keeping Found Money & Social Life, Social Cohesion \& Trust \\ 
pn61 & 0.00 & 0.00 & 0.50 & 0.50 & Values: Failing To Report Damage To Parked Vehicle & Social Life, Social Cohesion \& Trust \\ 
pp01 & 0.04 & 0.03 & 0.79 & 0.17 & Interest In Politics & Politics, Government \& Economy \\ 
pp02 & 0.08 & 0.05 & 0.59 & 0.35 & Overall Satisfaction With Democracy & Politics, Government \& Economy \\ 
pp03 & 0.08 & 0.05 & 0.56 & 0.40 & Feeling About Political Influence & Politics, Government \& Economy \\ 
pp04 & 0.08 & 0.06 & 0.63 & 0.31 & Trust In Federal Government & Politics, Government \& Economy \\ 
pp07 & 0.06 & 0.04 & 0.67 & 0.28 & Future: Active In Boycott & Politics, Government \& Economy \\ 
pp08 & 0.06 & 0.05 & 0.70 & 0.25 & Future: Active In Strike & Politics, Government \& Economy \\ 
pp09 & 0.06 & 0.04 & 0.72 & 0.23 & Future: Active In Demonstration & Politics, Government \& Economy \\ 
pp10 & 0.05 & 0.04 & 0.74 & 0.21 & Political Position: Left, Right & Politics, Government \& Economy \\ 
pp12 & 0.06 & 0.05 & 0.71 & 0.24 & Opinion On Swiss Army: Direction & Politics, Government \& Economy \\ 
pp13 & 0.09 & 0.06 & 0.55 & 0.40 & Opinion On Social Expenses: Direction & Politics, Government \& Economy \\ 
pp14 & 0.15 & 0.12 & 0.65 & 0.24 & Opinion On Joining The Eu: Direction & Politics, Government \& Economy \\ 
pp15 & 0.08 & 0.05 & 0.59 & 0.35 & Opinion On Chances For Foreigners: Direction & National Identity, Ethnicity \& Immigration \\ 
pp16 & 0.09 & 0.05 & 0.52 & 0.43 & Opinion On Environmental Protection: Direction & Environment \& Climate \\ 
pp17 & 0.10 & 0.05 & 0.51 & 0.43 & Opinion On Taxes On High Income: Direction & Politics, Government \& Economy \\ 
pp18 & 0.09 & 0.06 & 0.62 & 0.32 & Opinion On Nuclear Energy: Direction & Environment \& Climate \\ 
pp19\_1 & 0.08 & 0.05 & 0.55 & 0.40 & Party Support: Swiss Radical-Democratic Party & Politics, Government \& Economy \\ 
pp19\_3 & 0.08 & 0.05 & 0.66 & 0.28 & Party Support: Pss Swiss Socialist Party & Politics, Government \& Economy \\ 
pp19\_4 & 0.08 & 0.06 & 0.67 & 0.28 & Party Support: Udc Democratic Union Of The Centre & Politics, Government \& Economy \\ 
pp20 & 0.09 & 0.05 & 0.53 & 0.42 & Gender: Women In General Penalized & Gender \& Family \\ 
pp21 & 0.11 & 0.06 & 0.51 & 0.42 & Gender: Personnally Penalized & Gender \& Family \\ 
pp22 & 0.07 & 0.04 & 0.62 & 0.34 & Gender: In Favour Of Measures & Gender \& Family \\ 
pp23 & 0.15 & 0.07 & 0.38 & 0.55 & Gender: Personal Action & Gender \& Family \\ 
pp45 & 0.06 & 0.04 & 0.65 & 0.31 & General Trust In People & Social Life, Social Cohesion \& Trust \\ 
pp50 & 0.14 & 0.10 & 0.61 & 0.30 & Class Identification: Yes-No & Subjective Ses \\ 
pp51 & 0.11 & 0.10 & 0.79 & 0.11 & Class Identification (Spontanously): Which & Subjective Ses \\ 
pp52 & 0.17 & 0.15 & 0.70 & 0.15 & Class Identification (Non Spontaneously): Which & Subjective Ses \\ 
pp53 & 0.00 & 0.00 & 0.68 & 0.32 & Public Expenses: Protection Of The Environment & Politics, Government \& Economy \\ 
pp54 & 0.15 & 0.11 & 0.65 & 0.24 & Public Expenses: Health & Politics, Government \& Economy \\ 
pp55 & 0.09 & 0.07 & 0.72 & 0.21 & Public Expenses: Day Nursery & Politics, Government \& Economy \\ 
pp56 & 0.01 & 0.01 & 0.64 & 0.36 & Public Expenses: Police And Public Order & Politics, Government \& Economy \\ 
pp57 & 0.13 & 0.08 & 0.59 & 0.33 & Public Expenses: Research And Education & Politics, Government \& Economy \\ 
pp58 & 0.06 & 0.05 & 0.74 & 0.22 & Public Expenses: Military And Defense & Politics, Government \& Economy \\ 
pp59 & 0.09 & 0.06 & 0.63 & 0.31 & Public Expenses: Retirement & Politics, Government \& Economy \\ 
pp60 & 0.08 & 0.07 & 0.72 & 0.21 & Public Expenses: Culture And Art & Politics, Government \& Economy \\ 
pp61 & 0.03 & 0.02 & 0.62 & 0.35 & Public Expenses: Unemployment Benefit & Politics, Government \& Economy \\ 
pp62 & 0.11 & 0.09 & 0.69 & 0.22 & Public Expenses: Aid To Developing Countries & Politics, Government \& Economy \\ 
pp63 & 0.00 & 0.00 & 0.67 & 0.33 & Public Expenses: Social Aid & Politics, Government \& Economy \\ 
pp64 & 0.09 & 0.06 & 0.62 & 0.31 & Opinion On Joining The Eu: Joining Or Bilateral Agreements & Politics, Government \& Economy \\ 
pp65 & 0.18 & 0.15 & 0.70 & 0.14 & Opinion On Joining The Eu: Continuation Or Dissolution Of Bilateral Agreements & Politics, Government \& Economy \\ 
pp66 & 0.17 & 0.15 & 0.72 & 0.13 & Party Identification: Yes-No & Politics, Government \& Economy \\ 
pp67\_2 & 0.04 & 0.04 & 0.93 & 0.03 & Party Identification: Swiss Christian-Democrat Party & Politics, Government \& Economy \\ 
pp67\_3 & 0.06 & 0.06 & 0.90 & 0.04 & Party Identification: Pss Swiss Socialist Party & Politics, Government \& Economy \\ 
pp67\_4 & 0.04 & 0.04 & 0.93 & 0.03 & Party Identification: Udc Democratic Union Of The Centre & Politics, Government \& Economy \\ 
pp68 & 0.00 & 0.00 & 0.65 & 0.35 & Party Identification: Proximity & Politics, Government \& Economy \\ 
pp69 & 0.11 & 0.09 & 0.76 & 0.15 & Sympathy Political Party: Liberal Party & Politics, Government \& Economy \\ 
pp70 & 0.11 & 0.09 & 0.76 & 0.15 & Sympathy Political Party: Christian Democrats & Politics, Government \& Economy \\ 
pp71 & 0.08 & 0.08 & 0.83 & 0.09 & Sympathy Political Party: Social Democrats & Politics, Government \& Economy \\ 
pp72 & 0.09 & 0.08 & 0.85 & 0.07 & Sympathy Political Party: Swiss People's Party & Politics, Government \& Economy \\ 
pp73 & 0.09 & 0.08 & 0.80 & 0.12 & Sympathy Political Party: Green Party & Politics, Government \& Economy \\ 
pp74 & 0.07 & 0.05 & 0.64 & 0.31 & Opinion On Federalism: Direction & Politics, Government \& Economy \\ 
pp76 & 0.03 & 0.02 & 0.75 & 0.23 & Protection Of Environment: Importance & Environment \& Climate \\ 
pp77 & 0.13 & 0.10 & 0.65 & 0.26 & Protection Of Environment: Recycling & Environment \& Climate \\ 
pp78 & 0.21 & 0.19 & 0.69 & 0.13 & Protection Of Environment: Bin Bag Fees & Environment \& Climate \\ 
pp79 & 0.08 & 0.07 & 0.76 & 0.17 & Protection Of Environment: Organic And Ecological Products & Environment \& Climate \\ 
pp80 & 0.10 & 0.08 & 0.72 & 0.20 & Protection Of Environment: Purchase Of Fruits And Vegetables & Environment \& Climate \\ 
pr16 & 0.03 & 0.03 & 0.89 & 0.08 & Belief In God Or Something Divine & Religion \& Spirituality \\ 
pr20 & 0.11 & 0.10 & 0.81 & 0.09 & Feeling Of Spirituality & Religion \& Spirituality \\ 
pr21 & 0.05 & 0.05 & 0.84 & 0.11 & Feeling Of Religiosity & Religion \& Spirituality \\ 
pr23 & 0.11 & 0.10 & 0.79 & 0.12 & Attitude Towards Christianism & Religion \& Spirituality \\ 
pr24 & 0.13 & 0.12 & 0.76 & 0.12 & Attitude Towards Islam & Religion \& Spirituality \\ 
pr25 & 0.11 & 0.09 & 0.73 & 0.18 & Attitude Towards Hinduism & Religion \& Spirituality \\ 
pr26 & 0.12 & 0.10 & 0.77 & 0.13 & Attitude Towards Buddhism & Religion \& Spirituality \\ 
pr27 & 0.10 & 0.08 & 0.75 & 0.17 & Attitude Towards Judaism & Religion \& Spirituality \\ 
pr28 & 0.10 & 0.08 & 0.72 & 0.20 & Attitude Towards Atheism & Religion \& Spirituality \\ 
\midrule
\multicolumn{7}{l}{\textbf{SOEP}} \\ 
\midrule
plb0104\_h & 0.23 & 0.16 & 0.55 & 0.29 & Importance Of Employment & Occupation \& Education \\ 
plb0528\_h & 0.16 & 0.12 & 0.64 & 0.24 & Priority In Career Choice: Income & Occupation \& Education \\ 
plb0530\_h & 0.06 & 0.04 & 0.62 & 0.34 & Priority In Career Choice: Status & Occupation \& Education \\ 
plb0531\_h & 0.06 & 0.04 & 0.67 & 0.28 & Priority In Career Choice: Social Contact & Occupation \& Education \\ 
plb0532\_h & 0.12 & 0.08 & 0.60 & 0.31 & Priority In Career Choice: Compatibility Of Family And Work & Occupation \& Education \\ 
plc0115 & 0.10 & 0.06 & 0.58 & 0.35 & Social Security Contributions & Politics, Government \& Economy \\ 
pld0043 & 0.13 & 0.09 & 0.64 & 0.27 & Freq. Lend Friends Pers. Belongings & Social Life, Social Cohesion \& Trust \\ 
pld0044 & 0.16 & 0.12 & 0.64 & 0.24 & Freq. Lend Friends Money & Social Life, Social Cohesion \& Trust \\ 
pld0045 & 0.09 & 0.06 & 0.62 & 0.32 & Freq. Leave Door Unlocked & Social Life, Social Cohesion \& Trust \\ 
plh0001 & 0.17 & 0.13 & 0.63 & 0.24 & Democratic Ideal & Politics, Government \& Economy \\ 
plh0004 & 0.10 & 0.08 & 0.69 & 0.23 & Political Attitude & Politics, Government \& Economy \\ 
plh0007 & 0.05 & 0.03 & 0.69 & 0.28 & Interest In Politics & Politics, Government \& Economy \\ 
plh0011\_h & 0.08 & 0.05 & 0.53 & 0.42 & Party Preference: Yes & Politics, Government \& Economy \\ 
plh0012\_h1 & 0.05 & 0.04 & 0.81 & 0.15 & Party Preference: Spd & Politics, Government \& Economy \\ 
plh0012\_h2 & 0.04 & 0.04 & 0.80 & 0.16 & Party Preference: Cdu & Politics, Government \& Economy \\ 
plh0012\_h3 & 0.04 & 0.03 & 0.76 & 0.21 & Party Preference: Csu & Politics, Government \& Economy \\ 
plh0012\_h4 & 0.09 & 0.06 & 0.66 & 0.28 & Party Preference: Fdp & Politics, Government \& Economy \\ 
plh0012\_h5 & 0.06 & 0.05 & 0.76 & 0.19 & Party Preference: The Greens & Politics, Government \& Economy \\ 
plh0012\_h6 & 0.06 & 0.05 & 0.78 & 0.17 & Party Preference: The Left & Politics, Government \& Economy \\ 
plh0013\_h & 0.09 & 0.05 & 0.49 & 0.46 & Party Preference: Intensity & Politics, Government \& Economy \\ 
plh0014 & 0.24 & 0.17 & 0.54 & 0.29 & Government Interference & Politics, Government \& Economy \\ 
plh0015 & 0.00 & 0.00 & 0.46 & 0.54 & Extent And Type Of State Powers & Politics, Government \& Economy \\ 
plh0016 & 0.03 & 0.02 & 0.56 & 0.42 & Financial Security For Family & Politics, Government \& Economy \\ 
plh0017 & 0.03 & 0.02 & 0.53 & 0.45 & Care For Pre-School Children & Politics, Government \& Economy \\ 
plh0018 & 0.01 & 0.00 & 0.49 & 0.50 & Care For School Children & Politics, Government \& Economy \\ 
plh0019 & 0.14 & 0.09 & 0.55 & 0.36 & Financial Security When Unemployed & Politics, Government \& Economy \\ 
plh0020 & 0.00 & 0.00 & 0.59 & 0.41 & Job Creation Schemes & Politics, Government \& Economy \\ 
plh0021 & 0.15 & 0.10 & 0.57 & 0.34 & Financial Security When Sick & Politics, Government \& Economy \\ 
plh0022 & 0.08 & 0.05 & 0.51 & 0.44 & Help For The Sick & Politics, Government \& Economy \\ 
plh0023 & 0.19 & 0.13 & 0.57 & 0.30 & Financial Security In Old-Age & Politics, Government \& Economy \\ 
plh0024 & 0.05 & 0.03 & 0.52 & 0.45 & Help For The Elderly & Politics, Government \& Economy \\ 
plh0025 & 0.09 & 0.05 & 0.55 & 0.40 & Financial Security When Requiring Care & Politics, Government \& Economy \\ 
plh0026 & 0.07 & 0.04 & 0.53 & 0.43 & Help For Those Requiring Care & Politics, Government \& Economy \\ 
plh0030 & 0.16 & 0.10 & 0.51 & 0.39 & Worried About Childcare & Gender \& Family \\ 
plh0031 & 0.10 & 0.07 & 0.58 & 0.35 & Worried About Consequences Of Eu- Enlargement & Politics, Government \& Economy \\ 
plh0032 & 0.11 & 0.05 & 0.38 & 0.57 & Worried About Economic Development & Politics, Government \& Economy \\ 
plh0033 & 0.09 & 0.05 & 0.50 & 0.46 & Worried About Finances & Health \& Morale \\ 
plh0034 & 0.07 & 0.04 & 0.53 & 0.43 & Worried About Stability Of Financial Markets & Politics, Government \& Economy \\ 
plh0035 & 0.09 & 0.05 & 0.55 & 0.40 & Worried About Own Health & Health \& Morale \\ 
plh0036 & 0.11 & 0.05 & 0.45 & 0.49 & Worried About Environment & Environment \& Climate \\ 
plh0037 & 0.09 & 0.06 & 0.56 & 0.38 & Worried About Consequences From Climate Change & Environment \& Climate \\ 
plh0038 & 0.09 & 0.04 & 0.42 & 0.54 & Worried About Peace & Politics, Government \& Economy \\ 
plh0039 & 0.08 & 0.05 & 0.58 & 0.37 & Worried About Global Terrorism & Politics, Government \& Economy \\ 
plh0040 & 0.08 & 0.04 & 0.52 & 0.44 & Worried About Crime In Germany & Politics, Government \& Economy \\ 
plh0041 & 0.10 & 0.06 & 0.54 & 0.40 & Worried About Introduction Of Euro & Politics, Government \& Economy \\ 
plh0042 & 0.10 & 0.05 & 0.48 & 0.46 & Worried About Job Security & Politics, Government \& Economy \\ 
plh0054 & 0.07 & 0.04 & 0.52 & 0.44 & Peace And Order & Politics, Government \& Economy \\ 
plh0054\_1 & 0.12 & 0.06 & 0.48 & 0.46 & Political Goals: Prority Peace And Order & Politics, Government \& Economy \\ 
plh0056 & 0.07 & 0.04 & 0.48 & 0.48 & Citizen Influence & Politics, Government \& Economy \\ 
plh0056\_1 & 0.09 & 0.05 & 0.46 & 0.49 & Political Goals: Prority Citizen Influence & Politics, Government \& Economy \\ 
plh0058 & 0.05 & 0.02 & 0.47 & 0.50 & Fight Against Inflation & Politics, Government \& Economy \\ 
plh0058\_1 & 0.01 & 0.00 & 0.38 & 0.62 & Political Goals: Prority Rising Prices & Politics, Government \& Economy \\ 
plh0061 & 0.08 & 0.04 & 0.51 & 0.44 & Freedom Of Speech & Politics, Government \& Economy \\ 
plh0061\_1 & 0.15 & 0.08 & 0.48 & 0.44 & Political Goals: Prority Freedom Of Speech & Politics, Government \& Economy \\ 
plh0090 & 0.19 & 0.15 & 0.66 & 0.19 & Importance Of Circle Of Friends & Social Life, Social Cohesion \& Trust \\ 
plh0091\_v2 & 0.16 & 0.14 & 0.69 & 0.18 & Importance For Satisfaction: Work & Health \& Morale \\ 
plh0092\_v2 & 0.21 & 0.17 & 0.64 & 0.20 & Importance For Satisfaction: Family & Health \& Morale \\ 
plh0093\_v2 & 0.20 & 0.16 & 0.65 & 0.19 & Importance For Satisfaction: Friends & Health \& Morale \\ 
plh0094\_v2 & 0.21 & 0.17 & 0.64 & 0.19 & Importance For Satisfaction: Income & Health \& Morale \\ 
plh0095\_v2 & 0.24 & 0.19 & 0.59 & 0.23 & Importance For Satisfaction: Dwelling & Health \& Morale \\ 
plh0096\_v2 & 0.20 & 0.15 & 0.62 & 0.22 & Importance For Satisfaction: Political Influence & Health \& Morale \\ 
plh0097\_v2 & 0.14 & 0.12 & 0.74 & 0.14 & Importance For Satisfaction: Professional Success & Health \& Morale \\ 
plh0098\_v2 & 0.25 & 0.19 & 0.60 & 0.21 & Importance For Satisfaction: Leisure & Health \& Morale \\ 
plh0099\_v2 & 0.26 & 0.20 & 0.57 & 0.23 & Importance For Satisfaction: Health & Health \& Morale \\ 
plh0100\_v2 & 0.11 & 0.08 & 0.62 & 0.30 & Importance For Satisfaction: Protection Of Environment & Health \& Morale \\ 
plh0101 & 0.09 & 0.08 & 0.82 & 0.09 & Import.-Rlgion, Belief For Satisfaction & Health \& Morale \\ 
plh0104 & 0.08 & 0.05 & 0.57 & 0.38 & Importance: To Be Able To Afford Sth. & Subjective Ses \\ 
plh0105 & 0.07 & 0.04 & 0.54 & 0.42 & Importance: To Help Others & Social Life, Social Cohesion \& Trust \\ 
plh0106 & 0.09 & 0.06 & 0.59 & 0.35 & Importance: To Fulfil Once Potential & Health \& Morale \\ 
plh0107 & 0.13 & 0.09 & 0.62 & 0.29 & Importance: To Have Success In The Job & Occupation \& Education \\ 
plh0108 & 0.08 & 0.06 & 0.71 & 0.23 & Importance: To Have An Own House & Subjective Ses \\ 
plh0109 & 0.12 & 0.08 & 0.61 & 0.31 & Importance: To Have A Happy Marriage & Gender \& Family \\ 
plh0110 & 0.08 & 0.06 & 0.70 & 0.23 & Importance: To Have Children & Gender \& Family \\ 
plh0111 & 0.05 & 0.03 & 0.62 & 0.34 & Importance: To Be Socially And Politically Active & Social Life, Social Cohesion \& Trust \\ 
plh0112 & 0.09 & 0.07 & 0.64 & 0.29 & Importance: To Be Able To Travel And See The World & Health \& Morale \\ 
plh0113 & 0.24 & 0.20 & 0.62 & 0.19 & Entitlement Received Are Deserved & Politics, Government \& Economy \\ 
plh0114 & 0.27 & 0.19 & 0.53 & 0.28 & German Citizen Benefits Are Equitable & Politics, Government \& Economy \\ 
plh0152\_v1 & 0.09 & 0.06 & 0.60 & 0.33 & Satisfaction With Democracy [1987-1989] & Politics, Government \& Economy \\ 
plh0152\_v2 & 0.09 & 0.07 & 0.70 & 0.23 & Satisfaction With Democracy [2005,2010,2016] & Politics, Government \& Economy \\ 
plh0188 & 0.07 & 0.04 & 0.46 & 0.51 & Confident About Future & Health \& Morale \\ 
plh0192 & 0.02 & 0.02 & 0.61 & 0.37 & On The Whole Trust People & Social Life, Social Cohesion \& Trust \\ 
plh0193 & 0.05 & 0.03 & 0.60 & 0.37 & Nowadays Can't Trust Anyone & Social Life, Social Cohesion \& Trust \\ 
plh0194 & 0.11 & 0.07 & 0.57 & 0.35 & Caution Towards Foreigners & Social Life, Social Cohesion \& Trust \\ 
plh0195 & 0.08 & 0.05 & 0.61 & 0.34 & Most People Are Exploitive, Fair & Social Life, Social Cohesion \& Trust \\ 
plh0196 & 0.08 & 0.05 & 0.57 & 0.39 & Most People Are Helpful, Act In Own Interest & Social Life, Social Cohesion \& Trust \\ 
plh0197 & 0.06 & 0.04 & 0.71 & 0.24 & Willingness To Take Risks While Driving & Health \& Morale \\ 
plh0198 & 0.11 & 0.08 & 0.66 & 0.26 & Willingness To Take Risks In Financial Matters & Health \& Morale \\ 
plh0199 & 0.02 & 0.01 & 0.70 & 0.28 & Willingness To Take Risks In Leisure And Sports & Health \& Morale \\ 
plh0200 & 0.06 & 0.04 & 0.67 & 0.29 & Willingness To Take Risks In Occupation & Occupation \& Education \\ 
plh0201 & 0.18 & 0.14 & 0.64 & 0.22 & Willingness To Take Health Risks & Health \& Morale \\ 
plh0202 & 0.21 & 0.17 & 0.64 & 0.19 & Willingness To Take Risks In Trusting Other People & Social Life, Social Cohesion \& Trust \\ 
plh0204\_h & 0.06 & 0.04 & 0.58 & 0.38 & Self-Evaluation: Risk & Health \& Morale \\ 
plh0212 & 0.11 & 0.08 & 0.62 & 0.31 & Thorough Worker & Health \& Morale \\ 
plh0213 & 0.10 & 0.08 & 0.70 & 0.22 & Am Communicative & Health \& Morale \\ 
plh0214 & 0.12 & 0.09 & 0.63 & 0.28 & Am Sometimes Too Coarse With Others & Health \& Morale \\ 
plh0215 & 0.12 & 0.09 & 0.64 & 0.27 & Am Original & Health \& Morale \\ 
plh0216 & 0.07 & 0.05 & 0.64 & 0.31 & Worry A Lot & Health \& Morale \\ 
plh0217 & 0.11 & 0.08 & 0.58 & 0.35 & Able To Forgive & Health \& Morale \\ 
plh0218 & 0.08 & 0.06 & 0.65 & 0.29 & Tend To Be Lazy & Health \& Morale \\ 
plh0219 & 0.11 & 0.08 & 0.67 & 0.24 & Am Sociable & Health \& Morale \\ 
plh0220 & 0.09 & 0.07 & 0.70 & 0.23 & Value Artistic Experiences & Health \& Morale \\ 
plh0221 & 0.11 & 0.08 & 0.64 & 0.28 & Somewhat Nervous & Health \& Morale \\ 
plh0222 & 0.14 & 0.09 & 0.58 & 0.33 & Carry Out Tasks Efficiently & Health \& Morale \\ 
plh0223 & 0.12 & 0.08 & 0.63 & 0.29 & Reserved & Health \& Morale \\ 
plh0224 & 0.14 & 0.09 & 0.58 & 0.32 & Friendly With Others & Health \& Morale \\ 
plh0225 & 0.11 & 0.08 & 0.65 & 0.27 & Have Lively Imagination & Health \& Morale \\ 
plh0226 & 0.12 & 0.09 & 0.63 & 0.28 & Deal Well With Stress & Health \& Morale \\ 
plh0244 & 0.00 & 0.00 & 0.58 & 0.42 & Attitude Towards Future & Health \& Morale \\ 
plh0253 & 0.11 & 0.08 & 0.70 & 0.22 & Personal Patience & Health \& Morale \\ 
plh0254 & 0.10 & 0.07 & 0.65 & 0.28 & Personal Impulsivness & Health \& Morale \\ 
plh0255 & 0.14 & 0.11 & 0.67 & 0.22 & Inquisitive & Health \& Morale \\ 
plh0332 & 0.06 & 0.04 & 0.60 & 0.35 & Concerns Child Health & Health \& Morale \\ 
plh0335 & 0.10 & 0.07 & 0.65 & 0.28 & Worried About Own Retirement Pension & Politics, Government \& Economy \\ 
plh0336 & 0.17 & 0.10 & 0.51 & 0.39 & Worried About Cohesion In Society & Social Life, Social Cohesion \& Trust \\ 
plj0040 & 0.12 & 0.08 & 0.60 & 0.31 & Ties To Local Area & National Identity, Ethnicity \& Immigration \\ 
plj0043 & 0.09 & 0.07 & 0.70 & 0.23 & Ties To Local Area & National Identity, Ethnicity \& Immigration \\ 
plj0046 & 0.09 & 0.05 & 0.54 & 0.41 & Worried About Immigration To Germany & National Identity, Ethnicity \& Immigration \\ 
plj0047 & 0.10 & 0.05 & 0.44 & 0.51 & Worried About Hostility To Foreigners & National Identity, Ethnicity \& Immigration \\ 
plj0078 & 0.03 & 0.02 & 0.74 & 0.24 & Feel German & National Identity, Ethnicity \& Immigration \\ 
plj0081 & 0.14 & 0.09 & 0.59 & 0.32 & Connected With Country Of Origin & National Identity, Ethnicity \& Immigration \\ 
plj0082 & 0.08 & 0.03 & 0.41 & 0.55 & Feeling Of Not Belonging & National Identity, Ethnicity \& Immigration \\ 
plj0083 & 0.11 & 0.06 & 0.53 & 0.40 & Feel At Home In Country Of Origin & National Identity, Ethnicity \& Immigration \\ 
\midrule
\multicolumn{7}{l}{\textbf{UKHLS}} \\ 
\midrule
ahvwell & 0.07 & 0.05 & 0.62 & 0.33 & Importance Of Doing Well In Quals & Occupation \& Education \\ 
apvocimp & 0.18 & 0.14 & 0.61 & 0.25 & Importance Of Apprenticeship/Vocational & Occupation \& Education \\ 
carbuy1 & 0.20 & 0.12 & 0.51 & 0.37 & Comfort & Environment \& Climate \\ 
carbuy10 & 0.23 & 0.15 & 0.50 & 0.34 & Safety & Environment \& Climate \\ 
carbuy11 & 0.20 & 0.14 & 0.59 & 0.26 & Speed/Performance & Environment \& Climate \\ 
carbuy12 & 0.12 & 0.08 & 0.56 & 0.36 & Features - Sat-Nav, Cd Player, Music System, Power Steering Etc... & Environment \& Climate \\ 
carbuy2 & 0.18 & 0.10 & 0.47 & 0.43 & Costs - Purchase/Running/Resale Value/Tax/Insurance & Environment \& Climate \\ 
carbuy3 & 0.19 & 0.13 & 0.57 & 0.30 & Small Engine & Environment \& Climate \\ 
carbuy4 & 0.22 & 0.15 & 0.55 & 0.30 & Large Engine & Environment \& Climate \\ 
carbuy5 & 0.22 & 0.16 & 0.56 & 0.29 & Environmentally Friendly/Low Co2 Emissions & Environment \& Climate \\ 
carbuy6 & 0.25 & 0.11 & 0.34 & 0.55 & Electric Car/Van - One That's Plugged Directly Into An Electricity Supply & Environment \& Climate \\ 
carbuy7 & 0.17 & 0.12 & 0.59 & 0.29 & Style/Design/Image Of Brand/Model & Environment \& Climate \\ 
carbuy8 & 0.23 & 0.15 & 0.52 & 0.32 & Interior Space/Functionality/Boot Size & Environment \& Climate \\ 
carbuy9 & 0.17 & 0.09 & 0.44 & 0.47 & Reliability & Environment \& Climate \\ 
civicduty & 0.03 & 0.02 & 0.73 & 0.24 & Sense Of Civic Duty & Politics, Government \& Economy \\ 
colbens1 & 0.15 & 0.13 & 0.75 & 0.12 & Party Sympathy: Conservative & Politics, Government \& Economy \\ 
colbens2 & 0.15 & 0.13 & 0.74 & 0.13 & Party Sympathy: Labour & Politics, Government \& Economy \\ 
colbens3 & 0.17 & 0.14 & 0.68 & 0.18 & Party Sympathy: Liberal Democrats & Politics, Government \& Economy \\ 
demorient & 0.22 & 0.17 & 0.62 & 0.20 & Democratic Orientation & Politics, Government \& Economy \\ 
eumem & 0.12 & 0.12 & 0.85 & 0.03 & Should Uk Remain A Member Of The Eu & Politics, Government \& Economy \\ 
ocimpa & 0.05 & 0.04 & 0.62 & 0.34 & Occupation: Important: Job Security & Occupation \& Education \\ 
ocimpb & 0.08 & 0.06 & 0.69 & 0.26 & Occupation: Important: High Income & Occupation \& Education \\ 
ocimpe & 0.01 & 0.00 & 0.61 & 0.38 & Occupation: Important: Lot Leisure Time & Occupation \& Education \\ 
ocimpf & 0.00 & 0.00 & 0.63 & 0.37 & Occupation: Important: Interesting Job & Occupation \& Education \\ 
ocimpi & 0.06 & 0.04 & 0.69 & 0.27 & Occupation: Important: Contribute To Soc. & Occupation \& Education \\ 
ocimpk & 0.05 & 0.03 & 0.62 & 0.34 & Occupation: Important: Time For Family & Occupation \& Education \\ 
ocimpl & 0.08 & 0.06 & 0.71 & 0.23 & Occupation: Important: Job Help Others & Occupation \& Education \\ 
oprlg & 0.11 & 0.10 & 0.81 & 0.09 & Whether Belong To A Religion & Religion \& Spirituality \\ 
oprlg1\_1 & 0.01 & 0.01 & 0.89 & 0.10 & Religion: Church Of England/Anglican & Religion \& Spirituality \\ 
oprlg1\_11 & 0.01 & 0.01 & 0.98 & 0.01 & Religion: Muslim/Islam & Religion \& Spirituality \\ 
oprlg1\_2 & 0.00 & 0.00 & 0.95 & 0.05 & Religion: Roman Catholic & Religion \& Spirituality \\ 
oprlg3 & 0.07 & 0.06 & 0.84 & 0.10 & Religion Makes A Difference To Life & Religion \& Spirituality \\ 
scenv\_ftst & 0.21 & 0.17 & 0.61 & 0.22 & How Feel About Current Lifestyle And The Environment & Environment \& Climate \\ 
scenv\_grn & 0.00 & 0.00 & 0.65 & 0.35 & Being Green Is An Alternative Lifestyle & Environment \& Climate \\ 
scfannoy & 0.14 & 0.10 & 0.60 & 0.31 & Friends Get On My Nerves & Social Life, Social Cohesion \& Trust \\ 
scfcritic & 0.01 & 0.01 & 0.61 & 0.38 & Friends Criticise Me & Social Life, Social Cohesion \& Trust \\ 
scopecl200 & 0.06 & 0.03 & 0.52 & 0.45 & Affected By Climate Change Next 200 Years & Environment \& Climate \\ 
scopecl30 & 0.04 & 0.02 & 0.58 & 0.40 & Affected By Climate Change Next 30 Years & Environment \& Climate \\ 
scopfamb & 0.07 & 0.05 & 0.72 & 0.23 & Family Suffers If Mother Works Full-Time & Gender \& Family \\ 
scopfamd & 0.03 & 0.02 & 0.61 & 0.37 & Husband And Wife Should Contribute To Hh Income & Gender \& Family \\ 
scopfamf & 0.03 & 0.02 & 0.71 & 0.27 & Husband Should Earn, Wife Should Stay At Home & Gender \& Family \\ 
scopngbha & 0.19 & 0.15 & 0.64 & 0.22 & Belong To Neighbourhood & Social Life, Social Cohesion \& Trust \\ 
scopngbhb & 0.16 & 0.13 & 0.65 & 0.23 & Local Friends Mean A Lot & Social Life, Social Cohesion \& Trust \\ 
scopngbhc & 0.18 & 0.14 & 0.63 & 0.23 & Advice Obtainable Locally & Social Life, Social Cohesion \& Trust \\ 
scopngbhd & 0.17 & 0.13 & 0.63 & 0.24 & Can Borrow Things From Neighbours & Social Life, Social Cohesion \& Trust \\ 
scopngbhe & 0.11 & 0.07 & 0.57 & 0.36 & Willing To Improve Neighbourhood & Social Life, Social Cohesion \& Trust \\ 
scopngbhf & 0.20 & 0.15 & 0.60 & 0.24 & Plan To Stay In Neighbourhood & Social Life, Social Cohesion \& Trust \\ 
scopngbhg & 0.19 & 0.15 & 0.63 & 0.23 & Am Similar To Others In Neighbourhood & Social Life, Social Cohesion \& Trust \\ 
scopngbhh & 0.18 & 0.14 & 0.65 & 0.21 & Talk Regularly To Neighbours & Social Life, Social Cohesion \& Trust \\ 
scpundstnd & 0.09 & 0.06 & 0.65 & 0.28 & Spouse/Partner Understands Way I Feel & Gender \& Family \\ 
scrannoy & 0.20 & 0.16 & 0.63 & 0.21 & Family Gets On My Nerves & Gender \& Family \\ 
scrcritic & 0.05 & 0.04 & 0.63 & 0.34 & Family Criticises Me & Gender \& Family \\ 
scrletdwn & 0.21 & 0.15 & 0.59 & 0.26 & Family Lets Me Down & Gender \& Family \\ 
scrrely & 0.18 & 0.14 & 0.63 & 0.22 & Can Rely On Family & Gender \& Family \\ 
vote1 & 0.13 & 0.08 & 0.57 & 0.34 & Supports A Particular Political Party & Politics, Government \& Economy \\ 
vote2 & 0.15 & 0.09 & 0.51 & 0.40 & Closer To One Political Party Than Others & Politics, Government \& Economy \\ 
vote3\_1 & 0.20 & 0.13 & 0.53 & 0.34 & Party Support: Conservatives & Politics, Government \& Economy \\ 
vote3\_2 & 0.14 & 0.09 & 0.55 & 0.36 & Party Support: Labour & Politics, Government \& Economy \\ 
vote3\_3 & 0.22 & 0.13 & 0.48 & 0.39 & Party Support: Liberal Democrat & Politics, Government \& Economy \\ 
vote4\_1 & 0.05 & 0.05 & 0.88 & 0.07 & Party Affinity: Conservatives & Politics, Government \& Economy \\ 
vote4\_2 & 0.06 & 0.05 & 0.86 & 0.09 & Party Affinity: Labour & Politics, Government \& Economy \\ 
vote4\_3 & 0.12 & 0.10 & 0.73 & 0.18 & Party Affinity: Liberal Democrat & Politics, Government \& Economy \\ 
vote5 & 0.12 & 0.08 & 0.57 & 0.35 & Strength Of Support For Stated Party & Politics, Government \& Economy \\ 
vote6 & 0.06 & 0.04 & 0.73 & 0.23 & Level Of Interest In Politics & Politics, Government \& Economy \\ 
voteintent & 0.06 & 0.04 & 0.76 & 0.20 & Voting Intention & Politics, Government \& Economy \\ 
votenorm & 0.03 & 0.02 & 0.56 & 0.42 & Voting As A Social Norm & Politics, Government \& Economy \\ 
\bottomrule
\end{longtable}
\end{scriptsize}
```

\elandscape
\newpage

# Supplemental Materials B

Table B1 presents the full set of coefficients from an OLS model regressing $\omega$ on multiple covariates, described in Figure 2 in the main manuscript.

```{=latex}
\begin{table}[ht]
\begin{center}
\caption*{Table B1: Regression Model} 
\label{tab:sometab}
\begin{tabular}{l c c}
\hline
 & Estimates & SE \\
\hline
N Part. (log)                                       & $0.01$        
                                                    & $(0.01)$      \\
Duration                                            & $-0.00$ 
                                                    & $(0.00)^{***}$ \\
N Waves (Part.)                                     & $0.00$        
                                                    & $(0.00)$      \\
N Waves (Survey)                                    & $-0.00$       
                                                    & $(0.00)$      \\
Wave 1 (Decade)                                     & $-0.01$  
                                                    & $(0.00)^{**}$  \\
Panels: BHPS                                        & $0.08$        
                                                    & $(0.05)$      \\
Panels: GSS                                         & $0.04$        
                                                    & $(0.05)$      \\
Panels: HILDA                                       & $0.08$        
                                                    & $(0.05)$      \\
Panels: PSID                                        & $0.09$    
                                                    & $(0.04)^{*}$  \\
Panels: SHP                                         & $0.06$        
                                                    & $(0.05)$      \\
Panels: SOEP                                        & $0.06$        
                                                    & $(0.05)$      \\
Panels: UKHLS                                       & $0.09$        
                                                    & $(0.06)$      \\
Topics: Environment and Climate                     & $0.04$   
                                                    & $(0.01)^{**}$      \\
Topics: Gender and Family                           & $0.00$        
                                                    & $(0.01)$      \\
Topics: Health and Morale                           & $0.03$   
                                                    & $(0.01)^{**}$      \\
Topics: National Identity, Ethnicity, Immigration   & $0.02$        
                                                    & $(0.01)$      \\
Topics: Occupation and Education                    & $-0.02$       
                                                    & $(0.02)$      \\
Topics: Religion and Spirituality                   & $-0.00$       
                                                    & $(0.01)$      \\
Topics: Social Life, Social Cohesion and Trust      & $0.03$  
                                                    & $(0.01)^{***}$      \\
Topics: Subjective SES                              & $0.03$    
                                                    & $(0.01)^{*}$      \\
\hline
Adjusted R$^2$                                      & $0.77$     &   \\
N                                                   & $609$      &  \\
\hline
\multicolumn{2}{l}{\scriptsize{$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$}}
\end{tabular}
\end{center}
\end{table}
```

\newpage

# Supplemental Materials C

Figure C1 presents the alternative strategy we used to understand the effects of duration on intraindividual change. First, we restricted the sample such that we have at least 4 waves of observations for all participants. This reduced the number of items to 249. We then fitted the Life Course Adaption models to this sample. Second, we removed the last wave from all the observations, and refit the same models once more. In the final step, we calculated the $\omega$ values for each item. The Figure C1 shows a scatterplot for these two sets of observations.

```{=latex}
\begin{figure}[ht]
\begin{center}
\caption*{Figure C1: Difference in $\omega$ Across Two Sample Specifications} 
```
```{r fig3, out.width = "5in"}
knitr::include_graphics("../figures/figure_S3.png")
```
```{=latex}
\end{center}
\end{figure}
```