### 2023-09-29 manuscript/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
  \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{setspace}\doublespacing \usepackage{endnotes} \renewcommand{\enotesize}{\normalsize} \let\footnote=\endnote
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Quantifying Stability and Change in Personal Culture Using Panel Data},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Quantifying Stability and Change in Personal Culture Using Panel
Data}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle
\begin{abstract}
Recent work has produced conflicting interpretations on the question of
whether adults undergo durable changes in their personal culture over
time. This paper asserts that this divergence is less due to
researchers' findings than the approach taken -- a tournament of models
focusing on the presence or absence of change. To move beyond this, we
reanalyze the data used previously in this debate with an approach that
quantifies the processes of cultural difference. In doing so, we compare
how much cultural difference is explained by intrapersonal change and
interpersonal differences. Our results harmonize recent findings by
showing that, while most measures of personal culture show durable
change over time, the variance explained by intrapersonal change is
typically small compared to that of interpersonal differences at
baseline. Quantifying these processes offers new ways to address
theoretical questions of when and how their relative importance shifts.
We exemplify this approach by exploring how the importance of these
processes differs between college graduates and the rest of the
population.
\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Does personal culture---an individual's attitudes, beliefs, values, and
practices---change over the life course, or is it largely fixed by
adulthood? This question underlies an important contemporary debate in
sociology (Kiley and Vaisey 2020; Lersch 2023; Lizardo 2017) and has
deep roots in seemingly contradictory theoretical perspectives. For
example, pragmatist theories of action claim that changes in social
environments cause people to adapt their views and make new cultural
meanings (Gross 2009; Swidler 2001), while Bourdieusian practice
theories argue that the ``past conditions of production'' leave a mark
on people's personal culture that lasts throughout their lives (Bourdieu
1990). Models of social influence assume that people adapt their culture
in the face of new information (Goldberg and Stein 2018), while the
emphasis on cohort effects in models of aggregate social and cultural
change requires them to be open to change while young but become fairly
resistant to it as they age (Ryder 1965). Finally, life course theories
posit important changes over time as people advance through important
transitions in their lives (Bardi et al. 2009; Elder, Johnson, and
Crosnoe 2003).

Because all of these theoretical perspectives have some empirical
support, the current debate is not about whether their posited processes
exist at all, but about their relative contribution to explaining the
cultural differences we see in the world. That said, scholars have
struggled to reach a consensus on the importance of change during
adulthood. Over the span of a few years, most apparent changes in
people's responses to personal survey items appear to be transitory,
with little evidence of persistent change among adults (Kiley and Vaisey
2020; Vaisey and Kiley 2021). This suggests that change during adulthood
is not a major factor in explaining personal cultural differences. When
we consider a longer time horizon, however, there is evidence that
adults make at least some persistent changes (Lersch 2023).

These seemingly inconsistent findings are partly due to the fact that
researchers have taken a ``tournament of models'' approach to adjudicate
different theoretical perspectives (Lersch 2023: p.~228). That is,
researchers have compared the fit of different statistical models to
each item to determine whether it is more consistent with ``no change''
or ``any change.'' The size of these piles then serves as the primary
evidence for the truth of the theory (Kiley and Vaisey 2020; Lersch
2023; Vaisey and Kiley 2021).

We argue, however, that asking \emph{whether} people change on an item
may cause more problems than it solves. Being able to detect \emph{any}
change on an item depends not only on sample populations, measurement,
statistical methods, and power, but also on definitions of what counts
as ``change.'' These differences can limit the possibility of agreement
on which ``pile'' a variable belongs. Moreover, asking whether there is
evidence of change reduces to, ``Is there \emph{any} evidence of change
in this item?'' But such binary inquiries are ill-suited to address the
core of the theoretical debate: determining \emph{how much} the process
of intrapersonal change contributes to cultural differences.

There is enough evidence to assert that neither pre-adult socialization
nor the contemporary social context alone can explain the full range of
cultural differences. The question that remains is about the
\emph{relative} contributions of these processes. In the current debate,
for a single construct (e.g., support for gay rights), existing
approaches can only make claims about \emph{whether} intrapersonal
change is happening, not about how much. This also prevents
investigating under what conditions, across which groups, and in what
domains do we see differences in the relative importance of
intrapersonal change and pre-existing interpersonal differences. These
questions are about \emph{degree}, not existence; to take this debate
forward therefore requires precise measurement rather than declarations
of victory for a particular perspective.

In this paper, we move beyond the ``tournament of models'' approach and
introduce a method for quantifying the relative contributions of
interpersonal differences and intrapersonal change for a single item.
Drawing on seven panel data sets from five countries, we quantify the
amount of variance in survey responses that is due to between-person
differences at a single point during the survey versus the amount of
variance in responses that is due to systematic intrapersonal change
over the study period. This measure quantifies the amount of change over
the duration of the panel relative to the stable differences that exist
between people.

Using this measure, we observe similar patterns across the datasets that
have previously yielded conflicting interpretations. In all
datasets---despite their varied social contexts and duration---we find a
similar distribution of intrapersonal change. In general, we find that
intrapersonal change accounts for a substantially smaller amount of
variance in personal culture survey item than stable interpersonal
differences. Of course, whether the amount of intrapersonal change in an
item is large enough to be ``meaningful'' is a substantive and
theoretical question. Nevertheless, on many measures of personal culture
(some observed for over a decade), we do not observe enough
intrapersonal change for it to play a substantial role in explaining the
differences we observe between adults. Notable, interesting, and
important exceptions exist, of course, but they are departures from the
overall pattern.

In addition to advancing a theoretical debate, our method can also help
address substantive questions. To illustrate how it might be used by a
researcher with a specific topical interest, we investigate differences
in the relative importance of intrapersonal change for people with and
without a college degree. We find a smaller amount of systematic change
among college-educated respondents, suggesting that college solidifies
one's personal culture rather than fostering an openness to new
information. This offers a novel empirical basis to theorize about the
role of university education in shaping personal culture.

Taken together, the goal of this paper is to advance---and, hopefully,
transcend---debates about \emph{whether} adults change. People do
change, at least a little, on most things. Our method can help quantify
exactly \emph{how much}.

\hypertarget{background}{%
\section{Background}\label{background}}

\hypertarget{change-and-stability-in-personal-culture}{%
\subsection{Change and Stability in Personal
Culture}\label{change-and-stability-in-personal-culture}}

Recent debates about whether adults undergo intrapersonal cultural
change emerged in part because theories of cultural change at the
aggregate level tend to implicitly invoke one of two models of
individual behavior. The first, what Kiley and Vaisey (2020) call a
``Settled Dispositions Model'' (SDM), assumes that peoples' personal
culture is relatively fixed by the time they are adults. While they
might make temporary changes in their declarative culture in reaction to
their environments, this model assumes that people return to a settled
baseline over a short period of time. This model underlies theories of
cultural change that suggest people are imprinted by early socialization
experiences such as the ``past conditions of production'' a Bourdieusian
practice theory (Bourdieu 1990), cohort replacement theories of
aggregate change (Mannheim 1952; Ryder 1965), and control theories in
social psychology (Robinson 2007; Smith-Lovin and Heise 1988).

The second model summarized by Kiley and Vaisey (2020), an ``Active
Updating Model'' (AUM), posits that people continually update their
personal culture as they move through life. This model suggests people
change their personal culture as they adapt and make new meanings when
encountering new social environments, discourses, and information (Gross
2009; Swidler 2001). This model underlies, among others, theories of
cultural diffusion (Christakis and Fowler 2010), attitude alignment
(DellaPosta 2020), and polarization (Bail et al. 2018). It is also
implicit in most studies that ask whether specific experiences, changes
in social roles, or political events, affect personal culture (Gelman
and Margalit 2021; Slothuus and Bisgaard 2021; Visser and Mirabile
2004).

There is no reason to believe that only one of these two models is
``correct'' at all times and in all places. A population observed over
some period of time contains a mix of people who are noticeably changing
and people who are not. Instead, different perspectives argue that each
of these ideal-typical models is more operative at different times, for
different people, and for different elements of personal culture. For
example, adolescence and early adulthood is typically viewed as a
``formative period'' for personal cultural development and thus
characterized by higher rates of active updating, while middle age and
later life are potentially characterized more by settled dispositions
(Alwin and Krosnick 1991; Eaton et al. 2009; Krosnick and Alwin 1989).
Similarly, salient issues, such as views around gay rights in the 2010s;
issues that see substantial elite realignments, such as views around the
Vietnam conflict in the 1970s (Zaller 1992); and novel issues of public
opinion, such as views around vaccines during the Covid-19 pandemic
(Scoville et al. 2022), might be characterized by active updating, while
established issues of low salience are characterized by stability.

Empirically comparing which of these models better fit a broad range of
questions from the General Social Survey's rotating panels, Kiley and
Vaisey (2020) found limited evidence of durable change. While the
majority of items preferred the AUM, the amount of durable change
detected on these items was small. A substantial minority of questions
(39 percent) favored the SDM, meaning they showed no evidence of durable
change. Questions with more evidence of durable change included salient
issues like gay marriage and questions tapping ``public'' statements or
behaviors such as partisan identification and religious service
attendance. There was also more evidence of durable change among early
adults (people ages 18-30) than the rest of the population. Overall, the
researchers concluded that ``results ultimately suggest that real,
persistent attitude change is an uncommon phenomenon among adults'\,'
(Kiley and Vaisey 2020: p.~500; see also Vaisey and Kiley 2021). This
lack of durable change is consistent with other recent findings that
cohort replacement plays a somewhat larger role than period effects in
explaining differences in personal culture (Vaisey and Lizardo 2016).

On the other hand, the claim that change is a relatively infrequent
phenomenon has been difficult to square with research identifying
durable change as a result of social experiences across a number of
cultural dimensions, such as morality (Broćić and Miles 2021), trust
(Mewes et al. 2021), and concerns about immigration (Kratz 2021).
Similarly, there are longitudinal studies showing that cues from
political elites can change an individual's position on specific policy
issues (Slothuus and Bisgaard 2021; Zaller 1992) and that changes in
their close contacts and acquaintances can change individuals' attitudes
on group-related politics (DellaPosta 2018; Gelman and Margalit 2021).
Given how often we observe people change their personal culture, it is
hard to accept that adults do not change.

Drawing on these findings, Lersch (2023) challenged the SDM and AUM as a
``needless dichotomy,'' proposing the ``Life Course Adaption Model''
(LCAM) as an alternative. This model draws on the life course
perspective to model personal culture as a different linear trajectory
over the duration of a panel for each respondent. In doing this, Lersch
rectified two shortcomings of the AUM and SDM. First, the AUM, as
described by Kiley and Vaisey, posits that changes follow a Markov
process where responses at time \(t\) are a function of responses at
time \(t-1\) but not earlier time points. However, earlier life
experiences can (directly and indirectly) mold personal culture when
transitioning to new social roles or into new environments, even if
their initial impact is delayed. For example, childhood events might
influence views on family structures later when individuals form their
own families. Therefore, the LCAM considers influences from earlier than
just \(t-1\) on responses at time \(t\).

Second, Kiley and Vaisey's analysis of the AUM and SDM is based on
three-wave panel data over four years, which might not be extensive
enough to adjudicate the two models. Lersch evaluates the LCAM against
the AUM and SDM using panel data spanning a wider duration (from 3 to 36
years) and more waves (3 to 18), offering a better chance to observe
durable change.

When Lersch compared the LCAM to the AUM and SDM on data from five
countries, 297 of the 428 questions he analyzed preferred the LCAM,
suggesting that we observe some linear change for adults on most items.
The SDM was preferred on 112 items, and the rest did not yield
conclusive results. No questions favored the AUM. He concludes that
``new experiences over the life course {[}\ldots{]} can persistently
move individuals' personal culture in novel directions'' (Lersch 2023:
p.~243-244).

Despite differences in how these researchers interpret their findings,
the empirical results are not far apart. Kiley and Vaisey (2020) found
that the majority of items they tested favored the AUM, meaning there
was evidence of durable change among adults, a pattern consistent with
Lersch's results. Also Lersch (2023) found evidence of durable change on
most items, but this change only accounted for a small proportion of
variance explained. On average, people changed only about .07 standard
deviations over 10 years. This is consistent with Kiley and Vaisey's
finding that durable change is often small. Moreover, a quarter of the
items Lersch studied still preferred the SDM, suggesting that even with
different assumptions assumptions about change and more extensive data,
many questions show no evidence of durable change. Consistent with this,
in studies showing evidence of durable change in specific measures of
personal culture, the change is often small. For example, Brocic and
Miles (2021) estimate that completing graduate degrees in humanities,
arts, and social sciences shift peoples' moral relativism only about .2
standard deviations on average compared to people with no college
degree, and this is the largest effect they identify. And studies of
aggregate change show that, even on items where cohort effects explain
more variance than period effects, there is always evidence that some
people change over time (Vaisey and Lizardo 2016). In other words,
despite different interpretations, the results of previous work are in
many ways highly consonant.

\hypertarget{quantification-not-adjudication}{%
\subsection{Quantification, Not
Adjudication}\label{quantification-not-adjudication}}

The current debate can be simplified to asking whether people
\emph{ever} change. This is rarely theoretically interesting since the
answer is almost always ``yes.'' Over sufficient time and in a large
enough sample, researchers will likely observe some evidence for durable
individual change. A lack of evidence for change may be due to a poor
survey question, a low resolution in response options, or that the
question simply was not asked for long enough or to enough people.
Conversely, finding evidence for some change tells us only that it was
``not zero.'' From that alone, we learn little about how much change has
happened in a population.

A more theoretically productive approach begins with a model assuming
that during an observed time period people might remain stable, might
change a little, or might undergo significant shifts in their personal
culture. Lersch's LCAM does this by modeling each individual as
following a personal linear trajectory. But rather than debating whether
this model fits a particular data set better than a model that assumes
that people never change, researchers should quantify these individual
changes and compare those to the differences that exist between people
at the panel's start. In other words, the focus should shift from
asking, ``do people change?'' to asking, ``what are the relative
contributions of change and initial differences for explaining variance
in personal culture?''

We propose an approach that offers such a quantification. Our measure
relates two metrics that each reflect a combination of relevant
theoretical processes. The first metric, interpersonal differences at
the start of the survey, reflects the accumulated experiences of people
prior to entering the panel survey that shape how they answer questions.
Lersch calls this ``early imprinting.'' While commonly associated with
experiences during a formative period that result in settled
dispositions, this variation could also reflect experiences that
happened at any stage in life as long as they predate the panel and
consistently affect subsequent responses. For instance, for those people
entering the panel post-retirement, this ``imprinting'' might reflect
this pivotal life transition. Consequently, this metric also reflects
variation in individuals' social roles or statuses at the start of the
panel that were important in shaping their dispositions.

The second metric, the amount of intrapersonal linear change people make
during the panel, captures durable changes in personal culture over
time. This captures the set of processes collectively called
``persistent change'' or ``adaption'' by Lersch and ``active updating''
by Kiley and Vaisey. Lersch attributes these changes to social triggers
such as moving into a new environment or adopting new social roles
(although he never measures these directly). Additionally, they might
reflect the diffusion of new cultural forms across social networks, cues
from political elites or otherwise culturally influential leaders, the
emergence of issues in politics or culture, or large-scale social
shifts.

A third metric, which we might call ``residual variance'' or
``fluctuation,'' accounts for the remainder of variance in peoples'
responses. These non-durable changes emerge for a variety of reasons.
For example, people might not have a clear disposition on a particular
item as it is asked. Instead, they might internalize a broad set of
considerations and construct an opinion in the context of the survey
interview, with different considerations coming to the forefront of
their cognition during each interview (Feldman and Zaller 1992;
Tourangeau, Rips, and Rasinski 2000; Zaller 1992). This variance can
also include measurement error, such as misinterpreted questions,
erroneous response selections, or responses getting coded incorrectly.
Although this third metric at least partially reflects important
processes of personal culture, it does not directly touch on the ongoing
debate outlined here. Because of this, we quantify this component but
focus principally on the other two.

At a theoretical level, the coexistence of the sets of processes
captured in the two main metrics, interpersonal differences and
intrapersonal change, is undeniable, and they are linked in many ways.
Any intrapersonal change during one's life will likely manifest as
interpersonal differences by the time people enter a panel survey.
Furthermore, unless people are entirely socialized early on and never
deviate from these dispositions, we expect observing intrapersonal
change in a segment of the population when surveyed over time.

Furthermore, these metrics are not necessarily useful in isolation or
without context. What is important for the theoretical debate is the
relative contribution of these two components in explaining cultural
differences in a population. In other words, when we look at a
population of adults over some time frame, is there enough intrapersonal
change relative to interpersonal difference to view it as an important
explanation for why people differ from each other as observed? Only
through quantifying the two processes we can get closer to a true answer
to this question.

Yet the true utility of the proposed quantification goes beyond settling
past debates; it allows researchers to ask questions that were difficult
to explore with previous approaches. For example, classifying survey
items based on whether they show durable change overlooks possibly
important differences among those that do. By quantifying the relative
contributions of interpersonal differences and intrapersonal change,
researchers can gauge their relative importance in explaining
differences among survey items, between groups within a given
population, across time, and across societies.

\hypertarget{expectations}{%
\subsection{Expectations}\label{expectations}}

By shifting the question and adjusting the approach, we expect
consistent patterns across datasets previously analysed in the debate.
First, because previous work found evidence of durable change on most
measures of personal culture, we expect that we will find some
intrapersonal change in most items. But because these earlier studies
also found many questions (between 25 and 35 percent) lacking evidence
of \emph{any} durable change, we expect considerable variation in the
variance accounted for by intrapersonal change, with some questions
indicating essentially no variance attributable to such change.

Second, we expect the variance attributable to intrapersonal change to
be modest compared to interpersonal differences at the start of the
survey. This expectation should be uncontroversial, regardless of the
theoretical process assumed to be most relevant for a particular
question. Panel surveys do not encompass individuals' entire lifetimes.
Thus, even if intrapersonal change during adulthood is the major driver
of cultural differences, the change detected in a panel covering 20
years probably will not fully account for the entire variance of
interpersonal differences.

Our expectations so far largely echo what was observed in previous
studies. However, theoretical perspectives differ in how much variance
they would expect to be driven by intrapersonal change, enabling
quantification to provide a new perspective on existing questions.
``Formative period'' and cohort theories, which emphasize early-life
socialization as a strong influence on adult personal culture, suggest
that the variance from intrapersonal change would still fall short of
explaining interpersonal differences, even when we account for the
duration of observation. Conversely, other theoretical perspectives
emphasizing the importance of people's contemporaneous social
environment for personal culture might expect intrapersonal change to
account for a large portion of the variance, regardless of the duration.

Third, we expect the variance explained by each process ---interpersonal
differences and intrapersonal change--- to depend on specific attributes
of the question and panel. First, the higher the response resolution for
a given question and the larger the sample, the greater the variance
explained by intrapersonal change should be, as smaller changes in
intrapersonal change will be easier to detect. Second, the LCAM and
various theoretical perspectives that underlie it posit that social
transitions drive changes in personal culture. Since the probability
that any person experiences such a transition goes up the longer we
observe them, we expect panel duration to be positively related to the
proportion of variation explained by intrapersonal change.

\hypertarget{analytic-strategy}{%
\section{Analytic Strategy}\label{analytic-strategy}}

\hypertarget{data}{%
\subsection{Data}\label{data}}

We use data from seven nationally representative panel surveys from
Australia, Germany, Great Britain, Switzerland, and the United States
(summarized in Table 1), combining all the data files used in previous
work (Kiley and Vaisey 2020; Lersch 2023).\footnote{For more information
  on these data sources, see (Goebel et al. 2019; Income Dynamics 2013;
  Smith et al. 2022; Summerfield et al. 2011; Taylor 1996; University of
  Essex and Research 2019; Voorpostel et al. 2016).} These studies cover
a long period of time (the range of surveys spans from 1968 to 2021),
with 609 personal culture items that capture attitudes, beliefs, values,
self-assessments, self-descriptions, and behaviors (Alwin 2007). We
restricted the sample such that individuals between the ages of 18 and
79 are included without further elimination, and in all surveys, we used
all possible cases for which respondents provided responses. In the end,
the analyses that follow rely on a cross-national sample with a
cross-domain set of items to capture a broad range of individual
personal culture. Supplemental Materials A documents the list of all
variables used in the upcoming analyses.

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-1}The Description of the Data Sources}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llcc}
\toprule
Country & Survey & Period & Outcomes\\
\midrule
Australia & Household, Income and Labor Dynamics in Australia (HILDA) & 2001-2021 & 30\\
Germany & Socio-Economic Panel Study (SOEP) & 1984-2020 & 122\\
Great Britian & British Household Panel Survey (BHPS) & 1991-2008 & 83\\
Great Britian & Understanding Society/UK Household Longitudinal Study (UKHLS) & 2009-2020 & 66\\
Switzerland & Swiss Household Panel (SHP) & 1999-2019 & 77\\
\addlinespace
United States & General Social Survey (GSS) & 2006-2012 & 183\\
United States & Panel Study of Income Dynamics (PSID) & 1968-2019 & 48\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{life-course-adaption-model}{%
\subsection{Life Course Adaption
Model}\label{life-course-adaption-model}}

We start with Lersch's (2023) Life Course Adaption Model, which
formalizes survey responses at time \(t\) as a function of
individual-level random intercepts and slopes for survey age, what is
often called a mixed-effects growth curve model. This model formally
assumes a set of propositions about change that reflect the theoretical
debate to this point. First, consistent with the settled disposition
model, it assumes that people start the survey with cultural
differences, modeled as random intercepts for each respondent. Second,
it assumes that people change over time, taking the form of random
slopes for each respondent as a linear function of time. Third, it
assumes that people deviate around this baseline randomly over time,
reflecting ``fluctuation'' or short-term non-persistent change.
Formally, this can be written as

\[
y_{it} = \beta_0 + \alpha_{0i} + (\beta_1 + \alpha_{1i}) \text{year}_{it} + \epsilon_{it}
\]

\[
\begin{aligned}
\alpha_{0i} &\sim \mathcal{N}(0,\tau^2_{0}) \\
\alpha_{1i} &\sim \mathcal{N}(0,\tau^2_{1}) \\
\epsilon_{it} &\sim \mathcal{N}(0,\sigma^2)
\end{aligned}
\]

where \(\beta_0\) is the average intercept, \(\alpha_{0i}\) is the
random intercept component for individuals, \(\beta_1\) is the average
yearly change in the outcome, \(\alpha_{1i}\) is the random slope
component for individuals, and \(\epsilon_{it}\) is the random error
term that captures transitory disturbances. The \(\alpha_{0i}\) and
\(\alpha_{1i}\) terms are also allowed to covary.

\hypertarget{analysis-steps}{%
\subsection{Analysis Steps}\label{analysis-steps}}

As a first step in our analysis, we fit this LCAM to each of the 609
measures of personal culture outlined above. We then use these models to
derive two measures of systematic variance that correspond to the two
collections of theoretical processes outlined above: interpersonal
differences at baseline and intrapersonal change over time.

We estimate these using a form of variance decomposition. First, to
measure stable interpersonal differences, we calculate the sum of the
squared residuals derived from assuming that each person spends the
entire survey at the estimated midpoint of their LCAM
trajectory.\footnote{We use the midpoint estimate, rather than the value
  at the first wave, because this is the best measure of ``baseline''
  available under the assumption that the person does not change. The
  first wave measure alone contains an unknown amount of measurement
  error and transient fluctuations.} We then divide this value by the
total sum of squares and subtract this number from from 1, producing a
value we denote as \(V(D)\). This is the proportion of total variance
attributable to interpersonal differences. (This number is equivalent to
the conditional \(R^2\) value generated by a model with only random
intercepts for all respondents.) Second, we calculate another sum of
squared residuals generated by allowing the predictions to vary by
estimated LCAM trajectory. We divide this number by the total sum of
squares and subtract it from 1. We then subtract \(V(D)\) from this
value. This gives us a value we denote as \(V(C)\), the additional
proportion of variance accounted for by intrapersonal change.

These two numbers together reflect the total systematic variance of the
LCAM that can be attributed to either differences between people or
linear intrapersonal change over the course of the panel.\footnote{This
  does not mean that other factors might not explain additional variance
  if they were included in the model. For example, peoples' responses
  about religious beliefs and participation might systematically vary by
  the day of the week or month in which they are surveyed. This
  variation is moved into the ``random'' or non-systematic component of
  the model for us.} Our principal outcome of interest is the proportion
of systematic variance accounted for by intrapersonal change, which we
calculate as

\[
\omega = \frac{V(C)}{V(C) + V(D)}
\]

We can think of \(\omega\) as a type of Intraclass Correlation
Coefficient focused only on the systematic components of the model while
ignoring residual variance that might be attributable to measurement
error or non-persisting change (see also the variance decomposition of
period and cohort by Vaisey and Lizardo 2016). The higher this
proportion is, the more systematic variance in responses is attributable
to intrapersonal change during the panel. From now on, we use \(\omega\)
as a placeholder for \emph{systematic variance attributable to
intrapersonal change}.

Because this quotient summarizes variance and therefore has no natural
referent, it is hard to say at which values we observe ``a lot'' of
intrapersonal change. But summarizing questions in this way allows us to
compare the relative importance of our two broad processes across a
range of questions that might have different levels of non-systematic
variance. This gives us a sense of the relative prevalence of these two
processes for each question, which is not achievable under the
tournament of models approach. This also allows us to compare which
kinds of questions show more or less intrapersonal change relative to
interpersonal differences. While the tournament of models approach
allowed researchers to say whether questions in general showed
\emph{any} evidence of updating, this approach allows us to make more
specific claims about the prevalence of these theoretical processes for
particular questions relative to each other.

As a second step in our analysis, we use linear regression to conduct a
meta-analysis, modeling \(\omega\) as a function of features of the
question and panel. This allows us to explore how these features are
associated with the degree of intrapersonal change. These covariates
include the number of response options (measured as 2, 3-5, and more
than 5), the specific survey, the log of the number of participants, the
number of waves observed, the time period (i.e., decade) the question
was first asked, and the total duration of time the question was asked.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{variance-decomposition}{%
\subsection{Variance Decomposition}\label{variance-decomposition}}

Figure 1 plots the proportion of systematic variance attributable to
interpersonal differences and intrapersonal change for each of the 609
questions, plotted separately by panel. To broadly summarize the
results, all panels show a similar range of the proportion of systematic
variance attributable to intrapersonal change. Across panels, the median
value of \(\omega\) is 0.094 (mean of 0.098), with an interquartile
range of 0.053 to 0.139. All panels include questions where \(\omega\)
is essentially 0, and all panels include questions with \(\omega\)
values greater than 0.20. The maximum \(\omega\) value is 0.26.

\begin{figure}[htp]
\begin{center}
\caption{Proportions of Systematic Variance in Personal Culture}

\includegraphics[width=1\linewidth]{../figures/figure_1}

\end{center}
\footnotesize{\textit{Notes:} The figure shows $\omega$ and 1 - $\omega$ as the proportion of systematic variance attributable to intrapersonal change and interpersonal differences. See Supplemental Materials A for the full set of item values.}
\end{figure}

Across all questions, interpersonal differences account for a much
larger share of the systematic variance in responses than intrapersonal
change. Again, this is to be expected. Interpersonal differences capture
not just pre-adult socialization, but all accumulated experiences up to
the start of the panel that might influence personal culture.

To the extent that there are differences across the panels, the PSID has
the highest \(\omega\) values with mean .120 and median .135. While we
cannot disentangle features of the sample from features of the questions
asked to each sample, the specific samples for many PSID questions have
lower average ages than those from other panels. To the extent that
younger respondents might be more likely to make durable changes of
opinion, these higher estimates of intrapersonal change might reflect
the distinct age profile of respondents in this sample. At the other
end, the GSS has the lowest range of \(\omega\) with mean .073 and
median .069. This potentially reflects the fact that the GSS observes
people for a shorter duration, on average, than the other panels. If,
consistent with life course adaption theories, people are more likely to
make significant cultural changes the longer we observe them, then
duration likely affects the range of \(\omega\) (a point we explore more
below). However, the GSS results are still consistent with results from
the other panels.

While there are some differences between panels, these differences are
small compared to the differences within panels. For about 6 percent of
items, \(\omega\) is greater than 0.20. These questions tend to ask
about objectively changing external referents (e.g., confidence in
specific government leaders or political parties), life satisfaction, or
current financial position. At the other end, questions about religious
identification, views on gender roles, and support for civil liberties
tend to have very low estimates of intrapersonal change.

In contrast to the tournament of models approach, quantifying change
this way allows us to explore variation in the relative importance of
intrapersonal change across questions that all show evidence of change.
For example, Kiley and Vaisey (2020) found that confidence in the press
and confidence in religious leaders were both characterized by active
updating. Our results show that intrapersonal change is much more
important for explaining variance in confidence in the press (0.164)
than confidence in religion (0.049), even though both are updating.

Supplemental Materials A presents the estimated proportion of variance
attributable to interpersonal differences and intrapersonal change, the
estimated values of \(\omega\), and the proportion of residual variance
for each question. Interpersonal differences are almost always the
largest component of the total variance and tend to account for between
55 and 70 percent of total variance, while intrapersonal change is
always the smallest, typically accounting for between 3 and 8 percent of
total variance. Residual variance tends to account for between 22 and 37
percent of variance, though on several questions residual variance is
greater than 50 percent. This might indicate survey items with low
reliability or ones that capture genuinely rapid fluctuations.

As we noted above, the substantive importance of a given amount of
intrapersonal change depends on a range of factors, including how long
the panel runs, whether assumptions about linear change hold, and
whether the panel is capturing a distinctly turbulent period or a
distinctly stable period for the relevant item. However, if we assume
that the period under observation is ``typical'' for a question---not a
time of extremely heightened (or lowered) sensitivity or change---then
it does not seem realistic for that intrapersonal change accounts for a
large share of the cultural differences we see in the world.

\hypertarget{meta-analysis}{%
\subsection{Meta-Analysis}\label{meta-analysis}}

Figure 2 plots the results from a linear regression of \(\omega\) as a
function of question, panel, and sample features. These models also
include fixed effects for panels and topics, so coefficients reflect the
association within a panel and topical domain.

\begin{figure}[htp]
\begin{center}
\caption{Regression Model Estimating $\omega$}

\includegraphics[width=350px]{../figures/figure_2}

\end{center}
\footnotesize{\textit{Notes:} The model estimates the proportion of systematic variance attributable to intrapersonal change. It includes the log of the number of participants, the number of waves the variable was asked (capped at t = 10 for each item), the date of the first wave (measured as the year minus 1968 divided by 10), number of waves the question was asked averaged across participants, duration in years per item averaged across participants, and panel and topic fixed effects. Coefficients are estimated from suppressed intercept model based on predictions at participants' wave mid-point. Survey indicators and item topics not shown. See Supplemental Materials B for the full set of coefficients.}
\end{figure}

Figure 2 shows that the more response options respondents are given and
the larger the sample, the larger the values of \(\omega\). We interpret
these coefficients as suggesting greater resolution on a question makes
it easier to detect and model change. The earlier a question was asked,
even net of how long it was been asked, the lower \(\omega\).

Although we did not state expectations for how question content would
relate to \(\omega\), some of the associations between question
structure, panel duration, and panels themselves might be driven by
differences in the topics addressed by each panel. To address this, we
followed Hout et al. (2016) and Lersch (2023) in coding each question as
falling into one of nine different content categories and included these
as indicator variables in the meta analysis. Supplemental Materials B
presents the estimates for these content indicators. They show some
associations with \(\omega\), with questions about subjective SES;
social life, social cohesion, and trust; environment and climate; and
health and morale showing larger \(\omega\) values than questions about
religion and spirituality; politics, government and the economy; and
gender and family.

The most notable result in the meta analysis is the negative coefficient
attached to the duration of years covered by the question. The longer a
question is observed, the less is the value of \(\omega\). Theories of
personal cultural change that link changes in personal culture to social
experiences, including the LCAM, suggest that the longer we observe
respondents, the more likely people are to undergo potentially
transformative experiences and therefore the more variance would be
attributable to intrapersonal change. Finding a negative coefficient
here seems to challenge to that assumption.

As a further test of this finding, we compared the \(\omega\) values
when using the full duration of a panel compared to when we dropped the
final wave and therefore reduced the total duration of observation for
the question. If the coefficient reflects a true negative effect of
duration on \(\omega\), we should see that same effect within questions.
Results from this analysis are presented in Supplemental Materials C and
contradict the coefficient from the regression model; with a few
exceptions, we found that the longer we observe the same question, the
higher \(\omega\) is. We interpret this combination of findings as
suggesting that the kinds of questions asked for a longer time period
tend to demonstrate less intrapersonal change than questions asked over
shorter periods, rather than a true function of time. This supports the
conclusion that the GSS shows less intrapersonal change because of its
shorter duration.

\hypertarget{college-and-change-in-political-culture}{%
\subsection{College and Change in Political
Culture}\label{college-and-change-in-political-culture}}

These results mostly re-frame and align previous findings, but the value
of our approach lies in its ability to extend the debate to a broader
set of theoretical questions. To show this potential, we turn to a
specific empirical example: the relative importance of intrapersonal
change and interpersonal difference for explaining cultural variation by
level of education.

Previous work has established a positive relationship between education
and attitude stability, especially on issues related to American
politics. This stability is often attributed to education facilitating
``chronic information'' -- a general understanding of and attention to
the domain of American politics, including the positions held by major
parties and political figures and how issues relate to one another at a
logical or socio-logical level (Alvarez and Brehm 2002; Boutyline and
Vaisey 2017; Zaller 1992). These perspectives argue that because college
graduates have more knowledge of American politics, they are better able
to consistently connect the considerations in their cognition with the
answer choices they are presented with in a survey.

This work has tended to focus on the fact that college-educated
Americans give responses that are less likely to be affected by
measurement error or short-term influences than the rest of the
population (Alwin 2007; Zaller 1992). But this focus on the
non-systematic or residual component of variance across groups lumps the
two systematic forms of difference together. That is, it obscures the
fact that the amount of interpersonal differences and intrapersonal
change might also differ across these groups.

There are theoretical reasons to believe that education might be
associated with either more or less intrapersonal change. On one hand,
because college graduates are more connected to mainstream discourse and
elite signals, they might be more likely to make durable changes in
response to the emergence of new information, new issues, or political
realignments, while those without chronic information might display more
variance around an unchanging baseline (Zaller 1992). Conversely, it
could be that the observed stability of the college educated reflects
the fact that they have already formed durable opinions and are
relatively closed off to new information. If this is true, college could
be understood as a formative experience that solidifies some dimensions
of personal culture. Perhaps for those who do not attend college, later
life experiences might prove more important in forming or changing
personal culture, as these experiences potentially provide information
that college-educated peers have already received.

To compare these competing propositions, we calculate \(\omega\) values
separately for people with at least a bachelor's degree and people with
less than a bachelor's degree at wave 1\footnote{A small number of
  respondents report different highest degrees at wave 1, wave 2, and
  wave 3. Some of this is due to measurement error, and some of it is
  due to a small number of people obtaining a higher degree during the
  four years of the panel. Estimating the panel with highest at wave 1
  or highest degree reported across the panel produces functionally
  identical results.} of the three General Social Survey's panels. We
focus on the GSS because it contains the largest number of questions
tapping general political dispositions, which is the domain where
education has proven particularly relevant for understanding attitude
stability. The GSS also covers a turbulent window of American politics
from 2006 to 2014. This window covers the start of the Great Recession,
debates about federal intervention in and regulation of Wall Street, the
election of Barack Obama as the first black U.S. president, debates
about the role of the federal government in the health care sector, the
emergence of the Tea Party, and political realignment and clarification
on the issue of gay marriage, among other topics.

Figure 3 plots the distribution of differences in \(\omega\) values
between people with at least a bachelor's degree and people with less
than a bachelor's degree at wave 1 of the panel for 183 GSS items.
Values greater than 0 indicate that intrapersonal change accounts for
more systematic variance among college graduates than among those
without a college degree, while values less than 0 indicate the
opposite.

\begin{figure}[hbt]
\begin{center}
\caption{Difference in $\omega$ Across College Graduates and Non-Graduates}

\includegraphics[width=450px]{../figures/figure_3}

\end{center}
\footnotesize{\textit{Notes:} The figure shows the difference in $\omega$ values across college graduates and college non-graduates. Values above (below) 0 means that those with college degree have higher (lower) variance of intrapersonal change. The dashed red line marks 0 difference.}
\end{figure}

There is a clear pattern in Figure 3: for more than 80 percent of these
GSS items, intrapersonal change is a larger component of systematic
variance for people without a college degree. While most of these
differences are small in absolute terms (less than 2 percentage points),
several are greater than 5 percentage points. Given the distribution
observed in Figure 1 showing that the systematic variance attributable
to intrapersonal change averages around 0.09, a 5 percentage point
difference between groups is quite substantial.

To more clearly illustrate some of these differences, we highlight eight
questions designed to tap general political dispositions: partisan
identification (Democrat vs.~Republican) and ideological identification
(liberal vs.~conservative) on seven-point scales; four questions about
the government's role in improving the condition of the poor, paying
people's medical bills, giving special treatment to black people, and
doing things that private businesses could do, measured on five-point
scales; a question about whether the government should do more to reduce
income differences, measured on a seven-point scale; and one question
about whether black people should be given preferences in hiring,
measured on a five-point scale. We present estimates of \(\omega\) for
these eight questions, for both education groups, in Figure 4.

\begin{figure}[hbt]
\begin{center}
\caption{Difference in $\omega$ Across College Graduates and Non-Graduates on Political Culture}

\includegraphics[width=450px]{../figures/figure_4}

\end{center}
\footnotesize{\textit{Notes:} $\omega$ values across college graduates and college non-graduates on 8 political culture items from the General Social Survey (2006-2014).}
\end{figure}

On all eight questions presented in Figure 4, \(\omega\) values are
smaller for people with a college degree, meaning intrapersonal change
is less common among them. There are also large differences in
\(\omega\) values across questions for both groups. For example, on the
question of whether the government should try to solve more problems or
leave those problems to be solved by private businesses (``government do
more or less''), less than one percent of the systematic variance is
attributable to intrapersonal change for both groups. In other words,
while people might vacillate on this question at random (37 percent of
variance is residual for this question), there is functionally no
evidence that people make systematic changes of opinion on this issue
during the GSS panel.

In contrast, partisan identification and political ideology both show
larger values of \(\omega\) than most other questions, as well as a
larger absolute difference by degree status. Compared to the other
questions, intrapersonal change plays a much larger role in accounting
for partisan identification and political ideology. And this is
particularly true for respondents who do not have a college degree; the
\(\omega\) value for ideological identification among non-college
educated respondents is almost four times that of college-educated
respondents.

It is worth pointing out that this meaningful difference in \(\omega\)
values across education groups and across questions would not have been
detectable using previous methods. For partisan identification and
political ideology, both college-educated respondents and people with
less than college degree would likely favor the AUM or LCAM over the SDM
because they these questions both show evidence of some members of the
population making \emph{some} intrapersonal change. In other words, the
tournament of models obscures the fact that the relative importance of
intrapersonal change differs across these two groups, that intrapersonal
change explains more systematic variance for ideological identification
and partisan identification for people without a college degree, and
that there appears to be more durable change on questions of affirmative
action than on questions of government aid to black Americans.

We believe these patterns shed new light on the mechanisms underlying
differences in attitudes and behavior across groups. Something about
college attendance seems to crystallize personal culture. While these
results should not be interpreted as causal effects of attending college
-- they are potentially confounded by age, social class, race, gender,
and other factors that explain selection into higher education -- they
open up a set of new questions and dynamics to explore.

\hypertarget{discussion-and-conclusion}{%
\section{Discussion and Conclusion}\label{discussion-and-conclusion}}

The objective of this paper is to intervene in the recent debate on
whether people change their personal culture -- their attitudes,
beliefs, values, practices, and dispositions -- as they move through
their adult life. Instead of falling back on a ``tournament of models''
approach (Lersch 2023: p.~228) to proclaim victory for a particular
answer to this question, our goal is to refocus the debate to the
theoretically more productive question of the relative importance of
intrapersonal change for explaining differences in the personal culture.
We propose a new approach that quantifies the amount of observed
systematic variance that is attributable to either interpersonal
differences at baseline or intrapersonal change over the duration of a
panel.

Applying our proposed measure, \(\omega\), to 609 survey items from all
of the panel datasets previously discussed in this debate revealed a
consistent pattern. Nearly all questions show evidence of people making
some durable, intrapersonal change over time. Some show notably high
amounts of intrapersonal change. For some questions about life
satisfaction or views on government benefits, it seems plausible that
differences in adult experiences predominantly account for observed
differences between people.

However, intrapersonal change is often substantially less pronounced
than interpersonal differences, accounting for less than 10 percent of
systematic variance on average across all datasets. On numerous
questions, like those on civil liberties, abortion, generalized trust,
and civic duty, the systematic variance attributable to intrapersonal
change is essentially zero. For these questions, it seems there is not
enough cultural change during adulthood to warrant attributing the
differences we observe to experiences and social transitions; instead,
the primary source of the observed differences appears to stem from
experiences in childhood, adolescence, and early adulthood.

More than adjudicating past positions in this debate, the approach we
presented in this paper will enable scholars to explore more precisely
the relative contributions of change and stability in explaining
variation in personal culture, as well as how these contributions might
differ across populations. To illustrate this, we showed how the amount
of cultural difference explained by intrapersonal change varies
substantially by survey item but also by individuals' characteristics
such as their education. While age and other factors related to college
completion might confound this pattern, it suggests that college
completion may crystallize personal culture to an extent that renders
later adult experiences on political dispositions less influential.

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

Although we believe it is a major advance on previous work, the approach
we outline here has important limitations. We allocate all systematic
variance to one of two sets of theoretical processes: intrapersonal
change and interpersonal differences at baseline. Our approach does not
quantify the proportion of people who ``change,'' nor can we be sure
that the amount of interpersonal change we detect is driven by many
people making small changes or a few people making large changes.

Our results and interpretations also hinge on how we have defined
change. As does the LCAM, our approach treats cultural trajectories as
varying linear slopes for each respondent, thus assuming that change is
a linear function of time. This assumption simplifies reality in which
change likely also takes non-linear and discontinuous forms. People
might jump from one ``stable'' disposition to another or experience a
``turning point'' in their lives that upsets an otherwise stable
trajectory. In parts, our assumption of linearity is a limitation of the
data, as most questions are only observed for three to six waves. Panels
with more waves might allow researchers to loosen this assumption to
test alternative, more flexible models of change.

Similarly, our approach assumes that durable change is unidirectional.
This is a sensible assumption on short panels where classifying change
that lasts less than two years as durable seems unreasonable.
Practically, it means that the variance produced by people making
durable changes in their cultural dispositions to then return to a
previous state later in life is classified as residual variance, rather
than intrapersonal change. Longer panels and more flexible definitions
of ``change'' might allow us to account for such trajectories.

Finally, we had chosen to examine the broadest array of measures of
personal culture available to us, ranging from religious beliefs and
core values, to policy preferences, and even to the importance of
different features when buying a new car across five countries
(Australia, Germany, Switzerland, the United Kingdom and the United
States). Nevertheless, our findings remain limited to the kinds of
questions that are asked in panel surveys and in the contexts they were
administered, reflecting issues of general (national) politics, gender
roles, immigration and race relations, and general well being. Although
we have no reason to believe results to be different, our findings do
not directly speak to other dimensions of culture such as artistic
tastes, leisure activities, and time use.

\hypertarget{implications-for-cultural-sociology}{%
\subsection{Implications for Cultural
Sociology}\label{implications-for-cultural-sociology}}

Despite these limitations, we believe our method and findings have
important implications for social science research. Sociologists
interested in understanding cultural differences have largely asked
about the \emph{existence} of cultural change in adults (e.g., Kiley and
Vaisey 2020; Lersch 2023; Vaisey and Kiley 2021). Yet this approach has
inadvertently limited the debate. In any population, \emph{some} degree
of adult cultural change is inevitable. Although ideal types like the
settled dispositions and active updating models are useful, no
theoretical perspective would expect either early life socialization or
adult intrapersonal change to be the sole source of one's personal
culture. Our results reinforce this point, showing the relevance of both
factors and allowing their precise quantification.

This and other recent findings (Quinn et al. 2023; Stewart and Berkman
2023) suggest that it is theoretically more productive to measure the
relative importance of these two components in concrete, substantive
cases. Drawing a unified conclusion from survey items measuring various
cultural forms on different scales and across different time frames is
challenging. Nevertheless, the general pattern suggests that, for most
items, intrapersonal change in adulthood is not the primary reason we
see the differences we see between people in the world.

We are not claiming that adults remain static or that their changes are
inconsequential. Even quantitatively minor shifts, such as a 2
percentage-point change in support for gay marriage, can have massive
ramifications. While the majority may remain consistent in their views,
understanding the underlying mechanisms of even such numerically minor
shifts remains a crucial task for the sociology of culture.

Our findings suggest that understanding variation in personal culture
requires examining the conditions and experiences of early life. While
sociological research often focuses on transitions between social roles,
changes in social networks, and the experience of organizational
environments, these factors seem to account for a smaller proportion of
adult differences than early-life experience. Simply put, we need more
work on early-life socialization (Guhin, Calarco, and Miller-Idriss
2020).

This conclusion aligns with a range of recent causal inference work
suggesting that selection effects, rather than treatment effects,
predominantly account for personal cultural differences among
individuals in varied social roles and positions (Campbell and Horowitz
2016; Wodtke 2018). While some people clearly change as they transition
into new roles or environments, this change seems to be insufficient in
magnitude and duration to explain what are often pronounced differences
among people in diverse roles. These and our results suggest that when
observing differences in personal culture across social roles, such as
parenthood, education, or professional authority (Longest, Hitlin, and
Vaisey 2013), or across occupations (Weeden and Grusky 2005), selection
likely plays a large role in explaining these difference, though
exceptions always exist.

Our analysis does not provide an answer as to why intrapersonal change
seems to have limited impact on understanding cultural differences among
adults. The situations that promote durable change in personal culture
might simply be rare during adulthood. Alternatively, it is possible
that adults do encounter opportunities, necessities, and incentives for
change, but their ability or willingness to change decreases. All the
more it is important to research when and how social situations can
provoke durable change in adults.

Related to this, our results regarding education and political views
suggest that the importance of processes that lead to such change can
vary by group. Aligning with life course theories, it seems the
significance of experiences for cultural change may be contingent on
other, prior experiences. For example, factors that shape cultural
dispositions may likely differ for college and non-college graduates;
the latter might be more profoundly influenced by mid-life experiences
than the former in this regard. In trying to understand differences in
personal culture, sociologists should therefore pay more attention to
the heterogeneous effects that various factors including social events,
encounters, and situations can have.

Our findings also have implications for understanding cultural change at
the aggregate level. Given that intrapersonal change accounts for a
relatively small amount of variance among adults, many forms of cultural
change at the aggregate level are necessarily more likely driven cohort
replacement than by contemporaneous social conditions (Underwoood et al.
2022; Vaisey and Lizardo 2016). This likely holds true for cultural
change at the macro societal level as well as the micro level such as
within organizations, political parties, and professions. Our findings
also indicate that the impact of formative experiences, as opposed to
contemporaneous social factors, on cultural change is contingent upon an
individuals' level of education. The weight of these two factors might
thus change as the prominence of education shifts in the life course. If
scholars find similar differences across other social categorizations
like class or race, it might necessitate a more thorough integration of
demographic processes into the analysis of cultural change at the
population level.

\hypertarget{implications-for-survey-research}{%
\subsection{Implications for Survey
Research}\label{implications-for-survey-research}}

Our findings underscore the value of extended panel surveys to advance
theories of culture. On average, non-systematic fluctuations in
responses account for more than four times as much variance as
intrapersonal change, measured as linear change. Differentiating between
the two is therefore crucial for understanding cultural differences and
cannot be done with cross-sectional data. Instead we should extend
panels beyond two waves to gain more leverage to understand when and in
what form personal culture changes.

A key insight from our meta-analysis is that an increase in response
resolution (the range of response options survey respondents are given)
correlates with an increased share of systematic variance explained by
intrapersonal change. This may reflect specific issues being assessed by
survey items using different scales. Alternatively, it might imply that
when intrapersonal changes occurs, they are often subtle with answers
moving from ``agree'' to ``strongly agree,'' rather than from ``strongly
agree'' to ``strongly disagree.'' Such subtle changes require finer
response options to be detected.

Perhaps most important, we need more panel studies on youth. It is very
likely that the majority of adult differences in personal culture are
rooted in different experiences before age 18. By empirically
constraining our analyses to adult experiences, social scientists may
inadvertently concentrate on topics and questions that, while important,
might not be able to help explain major cultural differences in adult
populations.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

We believe the approach we outlined here can push past the ``needless
dichotomy'' implicit in the question of whether people change or not
(Lersch 2023). Characterizing questions as displaying change or not can
only take researchers so far, but the question of whether some questions
demonstrate more change than others, or whether some groups are
characterized by more stability than others, has the potential to weigh
in on a broader range of theoretical debates. We hope researchers find
our approach useful as they investigate these questions.

\theendnotes

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-alvarez2002}{}}%
Alvarez, R. Michael, and John Brehm. 2002. \emph{Hard Choices, Easy
Answers: Values, Information, and {American} Public Opinion}. Princeton,
N.J: Princeton University Press.

\leavevmode\vadjust pre{\hypertarget{ref-alwin2007}{}}%
Alwin, Duane F. 2007. \emph{Margins of {Error}: {A} {Study} of
{Reliability} in {Survey} {Measurement}}. Hoboken, N.J.: John Wiley \&
Sons.

\leavevmode\vadjust pre{\hypertarget{ref-alwin1991}{}}%
Alwin, Duane F., and Jon A. Krosnick. 1991.
{``\href{https://www.jstor.org/stable/2781642}{Aging, {Cohorts}, and the
{Stability} of {Sociopolitical} {Orientations} {Over} the {Life}
{Span}}.''} \emph{American Journal of Sociology} 97(1):169--95.

\leavevmode\vadjust pre{\hypertarget{ref-bail2018}{}}%
Bail, Christopher A., Lisa P. Argyle, Taylor W. Brown, John P. Bumpus,
Haohan Chen, M. B. Fallin Hunzaker, Jaemin Lee, Marcus Mann, Friedolin
Merhout, and Alexander Volfovsky. 2018. {``Exposure to Opposing Views on
Social Media Can Increase Political Polarization.''} \emph{Proceedings
of the National Academy of Sciences} 115(37):9216--21. doi:
\href{https://doi.org/10.1073/pnas.1804840115}{10.1073/pnas.1804840115}.

\leavevmode\vadjust pre{\hypertarget{ref-bardi2009}{}}%
Bardi, Anat, Julie Anne Lee, Nadi Hofmann-Towfigh, and Geoffrey Soutar.
2009. {``The Structure of Intraindividual Value Change.''} \emph{Journal
of Personality and Social Psychology} 97(5):913--29.

\leavevmode\vadjust pre{\hypertarget{ref-bourdieu1990}{}}%
Bourdieu, Pierre. 1990. \emph{The Logic of Practice}. Stanford, Calif. :
Stanford University Press, 1990.

\leavevmode\vadjust pre{\hypertarget{ref-boutyline2017}{}}%
Boutyline, Andrei, and Stephen Vaisey. 2017. {``Belief {Network}
{Analysis}: {A} {Relational} {Approach} to {Understanding} the
{Structure} of {Attitudes}.''} \emph{American Journal of Sociology}
122(5):1371--1447. doi:
\href{https://doi.org/10.1086/691274}{10.1086/691274}.

\leavevmode\vadjust pre{\hypertarget{ref-brocic2021}{}}%
Broćić, Miloš, and Andrew Miles. 2021. {``College and the {`{Culture}
{War}'}: {Assessing} {Higher} {Education}'s {Influence} on {Moral}
{Attitudes}.''} \emph{American Sociological Review} 00031224211041094.
doi:
\href{https://doi.org/10.1177/00031224211041094}{10.1177/00031224211041094}.

\leavevmode\vadjust pre{\hypertarget{ref-campbell2016}{}}%
Campbell, Colin, and Jonathan Horowitz. 2016. {``Does {College}
{Influence} {Sociopolitical} {Attitudes}?''} \emph{Sociology of
Education} 89(1):40--58. doi:
\href{https://doi.org/10.1177/0038040715617224}{10.1177/0038040715617224}.

\leavevmode\vadjust pre{\hypertarget{ref-christakis2010}{}}%
Christakis, Nicholas, and James Fowler. 2010. \emph{Connected: {The}
{Amazing} {Power} of {Social} {Networks} and {How} {They} {Shape} {Our}
{Lives}}. HarperCollins Publishers.

\leavevmode\vadjust pre{\hypertarget{ref-dellaposta2018}{}}%
DellaPosta, Daniel. 2018. {``Gay {Acquaintanceship} and {Attitudes}
Toward {Homosexuality}: {A} {Conservative} {Test}.''} \emph{Socius}
4:2378023118798959. doi:
\href{https://doi.org/10.1177/2378023118798959}{10.1177/2378023118798959}.

\leavevmode\vadjust pre{\hypertarget{ref-dellaposta2020}{}}%
DellaPosta, Daniel. 2020. {``Pluralistic {Collapse}: {The} {`{Oil}
{Spill}'} {Model} of {Mass} {Opinion} {Polarization}.''} \emph{American
Sociological Review} 85(3):507--36. doi:
\href{https://doi.org/10.1177/0003122420922989}{10.1177/0003122420922989}.

\leavevmode\vadjust pre{\hypertarget{ref-eaton2009}{}}%
Eaton, Asia A., Penny S. Visser, Jon A. Krosnick, and Sowmya Anand.
2009. {``Social {Power} and {Attitude} {Strength} {Over} the {Life}
{Course}.''} \emph{Personality and Social Psychology Bulletin}
35(12):1646--60. doi:
\href{https://doi.org/10.1177/0146167209349114}{10.1177/0146167209349114}.

\leavevmode\vadjust pre{\hypertarget{ref-elder2003}{}}%
Elder, Glen H., Monica Kirkpatrick Johnson, and Robert Crosnoe. 2003.
{``\href{https://doi.org/10.1007/978-0-306-48247-2_1}{The {Emergence}
and {Development} of {Life} {Course} {Theory}}.''} Pp. 3--19 in
\emph{Handbook of the {Life} {Course}}, \emph{Handbooks of {Sociology}
and {Social} {Research}}, edited by J. T. Mortimer and M. J. Shanahan.
Boston, MA: Springer US.

\leavevmode\vadjust pre{\hypertarget{ref-feldman1992}{}}%
Feldman, Stanley, and John Zaller. 1992. {``The {Political} {Culture} of
{Ambivalence}: {Ideological} {Responses} to the {Welfare} {State}.''}
\emph{American Journal of Political Science} 36(1):268--307. doi:
\href{https://doi.org/10.2307/2111433}{10.2307/2111433}.

\leavevmode\vadjust pre{\hypertarget{ref-gelman2021}{}}%
Gelman, Andrew, and Yotam Margalit. 2021. {``Social Penumbras Predict
Political Attitudes.''} \emph{Proceedings of the National Academy of
Sciences} 118(6):e2019375118. doi:
\href{https://doi.org/10.1073/pnas.2019375118}{10.1073/pnas.2019375118}.

\leavevmode\vadjust pre{\hypertarget{ref-data_soep}{}}%
Goebel, Jan, Markus M. Grabka, Stefan Liebig, Martin Kroh, David
Richter, Carsten Schröder, and Jürgen Schupp. 2019. {``The German
Socio-Economic Panel (SOEP).''} \emph{Jahrb{ü}cher f{ü}r
National{ö}konomie Und Statistik} 239(2):345--60.

\leavevmode\vadjust pre{\hypertarget{ref-goldberg2018}{}}%
Goldberg, Amir, and Sarah K. Stein. 2018. {``Beyond {Social}
{Contagion}: {Associative} {Diffusion} and the {Emergence} of {Cultural}
{Variation}.''} \emph{American Sociological Review} 83(5):897--932. doi:
\href{https://doi.org/10.1177/0003122418797576}{10.1177/0003122418797576}.

\leavevmode\vadjust pre{\hypertarget{ref-gross2009}{}}%
Gross, Neil. 2009. {``A {Pragmatist} {Theory} of {Social}
{Mechanisms}.''} \emph{American Sociological Review} 74(3):358--79. doi:
\href{https://doi.org/10.1177/000312240907400302}{10.1177/000312240907400302}.

\leavevmode\vadjust pre{\hypertarget{ref-guhin2020}{}}%
Guhin, Jeffrey, Jessica McCrory Calarco, and Cynthia Miller-Idriss.
2020. \emph{Whatever {Happened} to {Socialization}?} \emph{preprint}.
SocArXiv. doi:
\href{https://doi.org/10.31235/osf.io/zp2wy}{10.31235/osf.io/zp2wy}.

\leavevmode\vadjust pre{\hypertarget{ref-hout2016}{}}%
Hout, Michael, and Orestes P. Hastings. 2016. {``Reliability of the
{Core} {Items} in the {General} {Social} {Survey}: {Estimates} from the
{Three}-{Wave} {Panels}, 2006--2014.''} \emph{Sociological Science}
3:971--1002. doi:
\href{https://doi.org/10.15195/v3.a43}{10.15195/v3.a43}.

\leavevmode\vadjust pre{\hypertarget{ref-data_psid}{}}%
Income Dynamics, Panel Study of. 2013. {``Public Use Dataset.''}
\emph{The Survey Research Center Institute for Social Research,
University of Michgan}.

\leavevmode\vadjust pre{\hypertarget{ref-kiley2020}{}}%
Kiley, Kevin, and Stephen Vaisey. 2020. {``Measuring {Stability} and
{Change} in {Personal} {Culture} {Using} {Panel} {Data}.''}
\emph{American Sociological Review} 85(3):477--506. doi:
\href{https://doi.org/10.1177/0003122420921538}{10.1177/0003122420921538}.

\leavevmode\vadjust pre{\hypertarget{ref-kratz2021}{}}%
Kratz, Fabian. 2021. {``Do Concerns about Immigration Change After
Adolescence? How Education and Critical Life Events Affect Concerns
about Immigration.''} \emph{European Sociological Review}
37(6):987--1003.

\leavevmode\vadjust pre{\hypertarget{ref-krosnick1989}{}}%
Krosnick, Jon, and Duane F. Alwin. 1989. {``Aging and {Susceptibility}
to {Attitude} {Change}.''} \emph{Journal of Personality and Social
Psychology} 57:416--25. doi:
\href{https://doi.org/10.1037//0022-3514.57.3.416}{10.1037//0022-3514.57.3.416}.

\leavevmode\vadjust pre{\hypertarget{ref-lersch2023}{}}%
Lersch, Philipp M. 2023. {``Change in {Personal} {Culture} over the
{Life} {Course}.''} \emph{American Sociological Review} 88(2):220--51.
doi:
\href{https://doi.org/10.1177/00031224231156456}{10.1177/00031224231156456}.

\leavevmode\vadjust pre{\hypertarget{ref-lizardo2017}{}}%
Lizardo, Omar. 2017. {``Improving {Cultural} {Analysis}: {Considering}
{Personal} {Culture} in Its {Declarative} and {Nondeclarative}
{Modes}.''} \emph{American Sociological Review} 82(1):88--115. doi:
\href{https://doi.org/10.1177/0003122416675175}{10.1177/0003122416675175}.

\leavevmode\vadjust pre{\hypertarget{ref-longest2013}{}}%
Longest, K. C., S. Hitlin, and S. Vaisey. 2013. {``Position and
{Disposition}: {The} {Contextual} {Development} of {Human} {Values}.''}
\emph{Social Forces} 91(4):1499--1528. doi:
\href{https://doi.org/10.1093/sf/sot045}{10.1093/sf/sot045}.

\leavevmode\vadjust pre{\hypertarget{ref-mannheim1952}{}}%
Mannheim, Karl. 1952. {``The {Problem} of {Generations}.''} Pp. 276--322
in \emph{Essays on the {Sociology} of {Knowledge}}. New York: Oxford
University Press.

\leavevmode\vadjust pre{\hypertarget{ref-mewes2021}{}}%
Mewes, Jan, Malcolm Fairbrother, Giuseppe Nicola Giordano, Cary Wu, and
Rima Wilkes. 2021. {``Experiences Matter: A Longitudinal Study of
Individual-Level Sources of Declining Social Trust in the United
States.''} \emph{Social Science Research} 95:102537.

\leavevmode\vadjust pre{\hypertarget{ref-quinn2023}{}}%
Quinn, Joseph M., Robert E. Freeland, Kimberly B. Rogers, Jesse Hoey,
and Lynn Smith-Lovin. 2023. {``How Cultural Meanings of Occupations in
the u.s. Changed During the Covid-19 Pandemic.''} \emph{American
Behavioral Scientist} 67(1):125--47.

\leavevmode\vadjust pre{\hypertarget{ref-robinson2007}{}}%
Robinson, Dawn T. 2007. {``Control {Theories} in {Sociology}.''}
\emph{Annual Review of Sociology} 33(1):157--74. doi:
\href{https://doi.org/10.1146/annurev.soc.32.061604.123110}{10.1146/annurev.soc.32.061604.123110}.

\leavevmode\vadjust pre{\hypertarget{ref-ryder1965}{}}%
Ryder, Norman B. 1965. {``The {Cohort} as a {Concept} in the {Study} of
{Social} {Change}.''} \emph{American Sociological Review} 30(6):843--61.
doi: \href{https://doi.org/10.2307/2090964}{10.2307/2090964}.

\leavevmode\vadjust pre{\hypertarget{ref-scoville2022}{}}%
Scoville, Caleb, Andrew McCumber, Razvan Amironesei, and June Jeon.
2022. {``Mask {Refusal} {Backlash}: {The} {Politicization} of {Face}
{Masks} in the {American} {Public} {Sphere} During the {Early} {Stages}
of the {COVID}-19 {Pandemic}.''} \emph{Socius: Sociological Research for
a Dynamic World} 8:22.

\leavevmode\vadjust pre{\hypertarget{ref-slothuus2021}{}}%
Slothuus, Rune, and Martin Bisgaard. 2021. {``How {Political} {Parties}
{Shape} {Public} {Opinion} in the {Real} {World}.''} \emph{American
Journal of Political Science} 65(4):896--911. doi:
\href{https://doi.org/10.1111/AJPS.12550}{10.1111/AJPS.12550}.

\leavevmode\vadjust pre{\hypertarget{ref-data_gss}{}}%
Smith, Tom W., Michael Davern, Jeremy Freese, and Stephen L. Morgan.
2022. {``General Social Surveys, 1972-2022: Cumulative Codebook /
Principal Investigator, Tom w. Smith; Co-Principal Investigators,
Michael Davern, Jeremy Freese and Stephen l. Morgan.''} \emph{Chicago:
NORC}.

\leavevmode\vadjust pre{\hypertarget{ref-smith-lovin1988}{}}%
Smith-Lovin, Lynn, and David R. Heise. 1988. \emph{Analyzing {Social}
{Interaction}: {Advances} in {Affect} {Control} {Theory}}. New York:
Gordon; Breach Sciences Publishers.

\leavevmode\vadjust pre{\hypertarget{ref-stewart2023}{}}%
Stewart, Evan, and Diane Berkman. 2023. {``Shifting or Settled? Tracking
Racial Animus During COVID-19.''} \emph{Social Psychology Quarterly}
86(3):312--33.

\leavevmode\vadjust pre{\hypertarget{ref-data_hilda}{}}%
Summerfield, Michelle, Ross Dunn, Simon Freidin, Markus Hahn, Peter
Ittak, Milica Kecmanovic, Ning Li, Ninette Macalalad, Nicole Watson,
Roger Wilkins, and others. 2011. {``HILDA User Manual -- Release 10.''}
\emph{Melbourne Institute of Applied Economic and Social Research,
University of Melbourne}.

\leavevmode\vadjust pre{\hypertarget{ref-swidler2001}{}}%
Swidler, Ann. 2001. \emph{Talk of {Love}: {How} {Culture} {Matters}}.
Chicago: University of Chicago Press.

\leavevmode\vadjust pre{\hypertarget{ref-data_bhps}{}}%
Taylor, Marcia Freed. 1996. {``British Household Panel Survey User
Manual: Volume a-Introduction, Technical Report and Appendices.''}

\leavevmode\vadjust pre{\hypertarget{ref-tourangeau2000}{}}%
Tourangeau, Roger, Lance J. Rips, and Kenneth A. Rasinski. 2000.
\emph{The Psychology of Survey Response}. Cambridge, U.K. ; New York:
Cambridge University Press.

\leavevmode\vadjust pre{\hypertarget{ref-underwood2022}{}}%
Underwoood, Ted, Kevin Kiley, Wenyi Shang, and Stephen Vaisey. 2022.
{``Cohort Succession Explains Most Change in Literary Culture.''}
\emph{Sociological Science} 9:184--205.

\leavevmode\vadjust pre{\hypertarget{ref-data_ukhls}{}}%
University of Essex, Institute for Social, and Economic Research. 2019.
{``Understanding Society: Waves 1--9, 2009--2018 and Harmonised BHPS:
Waves 1--18, 1991--2009.''} \emph{UK Data Service}.

\leavevmode\vadjust pre{\hypertarget{ref-vaisey2021}{}}%
Vaisey, Stephen, and Kevin Kiley. 2021. {``A {Model}-{Based} {Method}
for {Detecting} {Persistent} {Cultural} {Change} {Using} {Panel}
{Data}.''} \emph{Sociological Science} 8:83--95. doi:
\href{https://doi.org/10.15195/v8.a5}{10.15195/v8.a5}.

\leavevmode\vadjust pre{\hypertarget{ref-vaisey2016}{}}%
Vaisey, Stephen, and Omar Lizardo. 2016. {``Cultural {Fragmentation} or
{Acquired} {Dispositions}? {A} {New} {Approach} to {Accounting} for
{Patterns} of {Cultural} {Change}.''} \emph{Socius} 2:2378023116669726.
doi:
\href{https://doi.org/10.1177/2378023116669726}{10.1177/2378023116669726}.

\leavevmode\vadjust pre{\hypertarget{ref-visser2004}{}}%
Visser, Penny S., and Robert R. Mirabile. 2004. {``Attitudes in the
Social Context: {The} Impact of Social Network Composition on
Individual-Level Attitude Strength.''} \emph{Journal of Personality and
Social Psychology} 779--95.

\leavevmode\vadjust pre{\hypertarget{ref-data_shp}{}}%
Voorpostel, Marieke, Robin Tillmann, Florence Lebert, Ursina Kuhn,
Oliver Lipps, Valérie-Anne Ryser, and B. Wernli. 2016. {``Swiss
Household Panel User Guide (1999--2015).''} \emph{Lausanne: FORS}.

\leavevmode\vadjust pre{\hypertarget{ref-weeden2005}{}}%
Weeden, Kim A., and David B. Grusky. 2005. {``The {Case} for a {New}
{Class} {Map}.''} \emph{American Journal of Sociology} 111(1):141--212.
doi: \href{https://doi.org/10.1086/428815}{10.1086/428815}.

\leavevmode\vadjust pre{\hypertarget{ref-wodtke2018}{}}%
Wodtke, Geoffrey T. 2018. {``The {Effects} of {Education} on {Beliefs}
about {Racial} {Inequality}.''} \emph{Social Psychology Quarterly}
81(4):273--94. doi:
\href{https://doi.org/10.1177/0190272518804145}{10.1177/0190272518804145}.

\leavevmode\vadjust pre{\hypertarget{ref-zaller1992}{}}%
Zaller, John. 1992.
\emph{\href{https://doi.org/10.1017/CBO9780511818691}{The {Nature} and
{Origins} of {Mass} {Opinion}}}. Cambridge: Cambridge University Press.

\end{CSLReferences}

\end{document}